---
title: "Analysis"
author: "Melle Stephen Oudshoorn"
date: "`r Sys.Date()`"
output: html_document
---

# Setup and work

```{r setup chunk, echo = TRUE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)

set.seed(19)
```

This section loads the crucial packages and creates the stagnant variables that are used in the rest of the work

```{r load packages and key lists}
library(tidyverse)
library(modelsummary)
library(psych)
library(patchwork)
library(fixest)
library(MatchIt)
#pretty tables
library(knitr)
library(kableExtra)
library(scales)
library(car)
library(gt)

eu_countries <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czech Republic",
  "Denmark", "Estonia", "Finland", "France", "Germany", "Greece", "Hungary",
  "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Malta", "Netherlands",
  "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", "Spain", "Sweden"
)

high_emission_sectors <- c( #division of high emissions sectors
      "Oil, Gas & Consumable Fuels",
      "Energy Equipment & Services",
      "Metals & Mining", 
      "Chemicals",
      "Paper & Forest Products",
      "Construction Materials",
      "Electric Utilities",
      "Gas Utilities",
      "Water Utilities",
      "Multi-Utilities",
      "Independent Power and Renewable Electricity Producers"
    )

environmental_vars <- c(
  "nox_emissions",
  "sox_emissions",
  "hazardous_waste",
  "energy_use_total",
  "renewable_energy_use_ratio",
  "waste_recycled_to_total_waste",
  "water_discharged",
  "water_withdrawal_total",
  "environmental_restoration_initiatives",
  "policy_waste",
  "renewable_energy_use",
  "policy_resource_efficiency",
  "climate_policy_statement"
)

social_vars <- c(
  "number_of_employees",
  "women_employees",
  "women_managers",
  "training_hours_total",
  "health_safety_policy",
  "salary_gap",
  "gender_pay_gap_percentage",
  "human_rights_policy",
  "csr_sustainability_external_audit",
  "supply_chain_health_safety_improvements",
  "donations_total",
  "policy_board_diversity"
)

governance_vars <- c(
  "shareholders_score",
  "shareholder_approval_significant_transactions",
  "independent_board_members",
  "csr_sustainability_reporting",
  "csr_sustainability_committee",
  "corruption_oversight",
  "stakeholder_engagement",
  "audit_committee_independence",
  "policy_executive_compensation_esg_performance",
  "whistleblower_protection",
  "policy_bribery_and_corruption"
)

all_vars <- c(environmental_vars, social_vars, governance_vars)

y_var <- "greenwashing_esg"
```

This section reads the thesis dataset that has been made by the create_panel.R document.
After this, it filters the observations that do not have an ESG score, firms that are in the financial sector, and the years 2013 and 2016 (due to not having enough observations to say anything meaningful about the sample).
Then, it calculates a lot of variables useful for the analysis.

```{r setup the basics}
df_panel_og <- read.csv("data/CSRD_thesis_data/panel_final.csv", check.names = FALSE) %>% 
  filter(!is.na(esg_score)) %>%  #filter out NA's for the ESG score
  filter(`GICS Sector Name` != "Financials") %>% #filter out the financial sector due to unreliable emissions
  filter(year >= 2017, year <= 2025) %>% #2026 and 2014 were in the original dataset but too sparse (around 100 observations each that align with the restrictives)
  filter(!is.na(total_debt_actual)) %>% 
  mutate(
    leverage = total_debt_actual/market_capitalization, #create leverage variable
    roa = net_income_after_tax / total_assets, #create roa variable
    tobins_q = market_capitalization / total_assets, #create tobin q variable
    share_turnover = turnover / market_capitalization, #share turnover
    high_emission_sector = `GICS Industry Name` %in% high_emission_sectors, #defines high emission sectors
    low_emission_sector = !(`GICS Industry Name` %in% high_emission_sectors),
    is_listed_eu = `Country of Exchange` %in% eu_countries, #defines whether a company is listed in the eu
    #size standards
    has_high_employees = ifelse(is.na(number_of_employees), FALSE, number_of_employees > 500), #whether a company exceeds NFRD size threshold regarding employees
    has_high_assets = ifelse(is.na(total_assets), FALSE, total_assets > 20000000), #whether a company exceeds NFRD size threshold regarding assets
    has_high_turnover = ifelse(is.na(turnover), FALSE, turnover > 40000000), #whether a company exceeds NFRD size threshold regarding turnover
    nfrd_size = has_high_employees, #if they have larger than 500 employees they are exposed
    #eu domicile
    in_EU_EU_domicile = `Country of Incorporation` %in% eu_countries,
    UK_based = `Country of Incorporation` %in% "United Kingdom",
    #NFRD/CSRD exposed
    treatment_exposed = has_high_employees & (has_high_assets | has_high_turnover) &
  (is_listed_eu | in_EU_EU_domicile),  #is in the EU
    #policy shock
    post_csrd = year >= 2024,
    post_announcement = year >= 2022,
    placebo_post_2019 = year >= 2019,
    #emissions
    total_emissions_log = log(scope_1_and_scope_2_and_scope_3_combined_reported_or_estimated_total + 1),
    scope1_emissions_log = log(scope_1_combined_reported_or_estimated_total + 1),
    scope2_emissions_log = log(scope_2_combined_reported_or_estimated_total + 1),
    scope3_emissions_log = log(scope_3_combined_reported_or_estimated_total + 1),
    inv_emissions = -total_emissions_log, #inverted so that higher = better performance
    #scope1 reported or estimated
    scope1_reported = if_else(scope_1_combined_reported_or_estimated_method == "Reported_value", 1L, 0L,
      missing = NA_integer_),
    scope1_estimated = if_else(scope_1_combined_reported_or_estimated_method %in% c(
    "Aggregated_model","Winsorized_low","Extrapolated","Winsorized_high","Energy_extrapolated","Energy_model"),
    1L, 0L, missing = NA_integer_),
    #other pollutants
    nox_log = log(nox_emissions + 1),
    sox_log = log(sox_emissions + 1),
    waste_log = log(hazardous_waste + 1),
    #standardise emissions to match
    total_emissions_std = as.numeric(scale(total_emissions_log)),
    scope1_emissions_std = as.numeric(scale(scope1_emissions_log)),
    scope2_emissions_std = as.numeric(scale(scope2_emissions_log)),
    scope3_emissions_std = as.numeric(scale(scope3_emissions_log)),
    nox_std = as.numeric(scale(nox_log)),
    sox_std = as.numeric(scale(sox_log)),
    waste_std = as.numeric(scale(waste_log)),
    esg_std = as.numeric(scale(esg_score)),
    env_pillar_std = as.numeric(scale(environmental_pillar_score)),
    #group
    group = case_when(
      #treatment
      treatment_exposed == TRUE ~ "Treatment",
      #placebo
      UK_based == TRUE ~ "Placebo",
      #control (excl. UK)
      treatment_exposed == FALSE & UK_based == FALSE ~ "Control", # and UK_based == FALSE
      #fallback
      TRUE ~ "Other"
    )
  )

df_panel_og <- df_panel_og %>%
  group_by(`Identifier (RIC)`) %>%
  arrange(year, .by_group = TRUE) %>%
  mutate(
    emissions_change = ifelse(
      !is.na(lag(total_emissions_log)) & lag(total_emissions_log) != 0,
      (total_emissions_log - lag(total_emissions_log)) / lag(total_emissions_log),
      NA_real_
    )
  ) %>%
  mutate(
    emissions_trajectory = mean(emissions_change[year %in% 2019:2021], na.rm = TRUE),
    emissions_trajectory = replace_na(emissions_trajectory, 0)
  ) %>%
  ungroup()

consistency_check <- df_panel_og %>%
  filter(year < 2021) %>%
  group_by(`Identifier (RIC)`) %>%
  summarize(
    n_pre = n_distinct(year),
    .groups = "drop"
)

keep_ids <- consistency_check %>%
  filter(n_pre >= 3) %>% #AT LEAST THREE OBSERVATIONS PER FIRM, and has an observation in y
  pull(`Identifier (RIC)`)

df_panel <- df_panel_og %>%
  filter(`Identifier (RIC)` %in% keep_ids)
```

This chunk defines a function to construct the disclosure scores from the data, as provided by the LSEG database, after which it is applied to the dataset.

```{r build the disclosure score}
construct_disclosure_scores <- function(df, env_vars, soc_vars, gov_vars) {
  
  # Helper function to turn data into 0/1 points
  score_var <- function(x) {
    is_binary <- is.logical(x) || all(x %in% c(TRUE, FALSE, 1, 0, NA))
    if (is_binary) return(as.integer(!is.na(x) & as.logical(x)))
    return(as.integer(!is.na(x)))
  }

  # 1. Score the matrices
  # We use the filtered lists to ensure we only count 'responsive' variables
  env_scores <- df[env_vars] %>% mutate(across(everything(), score_var))
  soc_scores <- df[soc_vars] %>% mutate(across(everything(), score_var))
  gov_scores <- df[gov_vars] %>% mutate(across(everything(), score_var))

  # 2. Sum the scores
  # env_sum is now just the sum of the filtered environmental variables
  df <- df %>%
    mutate(
      environmental_disclosure_score = rowSums(env_scores),
      social_disclosure_score        = rowSums(soc_scores),
      governance_disclosure_score    = rowSums(gov_scores),
      # Combined ESG disclosure score
      esg_disclosure_score           = environmental_disclosure_score +
                                       social_disclosure_score +
                                       governance_disclosure_score
    )
  
  return(df)
}

# 3. Run the calculation
df_panel <- construct_disclosure_scores(
  df_panel,
  environmental_vars, # Use the cleaned list here
  social_vars,
  governance_vars
)
```

```{r describe the disclosure variables}
psych::describe(df_panel %>%
  select(
    esg_disclosure_score,
    environmental_disclosure_score,
    social_disclosure_score,
    governance_disclosure_score
  ))

df_esg_disclosure_long <- df_panel %>%
  select(
    esg_disclosure_score,
    environmental_disclosure_score,
    social_disclosure_score,
    governance_disclosure_score
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )

ggplot(df_esg_disclosure_long, aes(x = value)) +
  geom_histogram(bins = 50) +
  facet_wrap(~ variable, scales = "free") +
  labs(
    title = "Distribution of ESG Variables",
    x = NULL,
    y = "Count"
  ) +
  theme_minimal()
```

above, the disclosure scores are described. We observe a large influx at the environemntal disclosure score being 1, which skews the model quite a bit.
below, the greenwashing indeces are computed according to the main method composed by yu et al.

```{r computing the greenwashing index according to yu et al.}
df_panel <- df_panel %>%
  group_by(year, `GICS Sector Code`) %>%  #group by year and sector to remove mechanical industry variation
  mutate(
    #standardise the disclosure scores within industry-year groups
    esg_disclosure_z = as.numeric(scale(esg_disclosure_score)),
    environmental_disclosure_z = as.numeric(scale(environmental_disclosure_score)),
    social_disclosure_z = as.numeric(scale(social_disclosure_score)),
    governance_disclosure_z = as.numeric(scale(governance_disclosure_score)),
    #standardise performance scores
    esg_performance_z = as.numeric(scale(esg_score)),
    environmental_performance_z = as.numeric(scale(environmental_pillar_score)),
    social_performance_z = as.numeric(scale(social_pillar_score)),
    governance_performance_z = as.numeric(scale(governance_pillar_score)),
    
    #compute greenwashing indeces (standardised disclosure - standardised performance)
    
    #baseline greenwashing index
    greenwashing_esg = esg_disclosure_z - esg_performance_z,
    
    #pillared greenwashing index
    greenwashing_environmental = environmental_disclosure_z - environmental_performance_z,
    greenwashing_social = social_disclosure_z - social_performance_z,
    greenwashing_governance = governance_disclosure_z - governance_performance_z,
    
    #robustness check
    greenwashing_environmental_alt = esg_disclosure_z - environmental_performance_z #robustness check
  ) %>% 
  ungroup()

#full sample alternative (robustness)
df_panel <- df_panel %>%
  mutate(
    esg_disclosure_z_raw = as.numeric(scale(esg_disclosure_score)),
    esg_performance_z_raw = as.numeric(scale(esg_score)),
    greenwashing_esg_raw = esg_disclosure_z_raw - esg_performance_z_raw
  )

#year clustered alternative (robustness)
df_panel <- df_panel %>%
  group_by(year) %>%
  mutate(
    esg_disclosure_z_y = as.numeric(scale(esg_disclosure_score)),
    esg_performance_z_y = as.numeric(scale(esg_score)),
    greenwashing_esg_y = esg_disclosure_z_y - esg_performance_z_y
  ) %>%
  ungroup()

#industry clustered alternative (robustness)
df_panel <- df_panel %>% 
  group_by(`GICS Sector Code`) %>%
  mutate(
    esg_disclosure_z_i = as.numeric(scale(esg_disclosure_score)),
    esg_performance_z_i = as.numeric(scale(esg_score)),
    greenwashing_esg_i = esg_disclosure_z_i - esg_performance_z_i
  ) %>%
  ungroup()
  
```



```{r computing the greenwashing index according to Dorfleitner & Utz}
#esg vs total emissions and energy use
gw_model <- feols(esg_std ~ total_emissions_std + energy_use_total | `Identifier (RIC)` + year,
                  cluster = ~`Identifier (RIC)` + year,
                  data = df_panel_og)
#E-pillar of esg vs total ghg and energy use
gw_env_model <- feols(env_pillar_std ~ total_emissions_std + energy_use_total | `Identifier (RIC)` + year,
                       cluster = ~`Identifier (RIC)` + year,
                       data = df_panel_og)

#extracting the residuals safely

#compute the greenwashing as the residual
df_panel_og <- df_panel_og %>%
  mutate(
    # ESG-based residual GW
    gw_dorfleitner_utz = esg_std - predict(gw_model, newdata = .),
    
    # Environment-only residual GW
    gw_env_dorfleitner_utz = env_pillar_std - predict(gw_env_model, newdata = .)
  )
```

## Computing the matched DF's 

This section creates the dataframes that are also used for analysis with propensity score matching and coarsed exact matching.

```{r propensity score matching}
covariates <- c(
  "log_total_assets",
  "roa",
  "leverage",
  "tobins_q",
  "share_turnover",
  "gics_sector_code"
)

df_baseline <- df_panel %>%
  filter(year == 2019) %>% 
  mutate(
    log_total_assets = ifelse(
      !is.na(total_assets) & total_assets > 0,
      log(total_assets),
      NA_real_
    ),
    gics_sector_code = factor(`GICS Sector Code`)
  ) %>%
  # Only keep firms that have all must‑have covariates non‑missing
  filter(
    across(all_of(covariates), ~ !is.na(.x)),
    !is.na(treatment_exposed),
    !is.na(`Country of Incorporation`),
    !is.na(gics_sector_code)
  ) %>%
  distinct(`Identifier (RIC)`, .keep_all = TRUE)

psm_formula <- as.formula(
  paste("treatment_exposed ~",
        paste(covariates, collapse = " + "))
)

psm_match <- matchit(
  formula = psm_formula,
  data = df_baseline,
  method = "nearest",
  distance = "logit"
)

matched_ids <- match.data(psm_match) %>%
  pull(`Identifier (RIC)`) %>%
  unique()

df_psm_panel <- df_panel %>%
  filter(`Identifier (RIC)` %in% matched_ids)

summary(psm_match)
```

This yield a panel with 4577 matched observations

```{r mahalanobis matching}
mdm_match <- matchit(psm_formula,
                     data = df_baseline,
                     method = "nearest",
                     distance = "mahalanobis")

matched_ids_mdm <- match.data(mdm_match)$`Identifier (RIC)`
df_mdm_panel <- df_panel %>% filter(`Identifier (RIC)` %in% matched_ids_mdm)

summary(mdm_match)
```


```{r coarsened exact matching}
#match based on the formula
cem_match <- matchit(psm_formula, 
                     data = df_baseline, 
                     method = "cem",
                     cutpoints = list(
                       "log(total_assets)" = 15, # Break assets into 15 tight groups
                       "roa" = 10,               # Break ROA into 10 tight groups
                       "leverage" = 10           # Break leverage into 10 tight groups
                     ))
#get matched IDS
matched_ids_cem <- match.data(cem_match)$`Identifier (RIC)`
#filter the data
df_cem_panel <- df_panel %>% 
  filter(`Identifier (RIC)` %in% match.data(cem_match)$`Identifier (RIC)`)

summary(cem_match)
```

This yields a panel with 14,411 matched observations.

After, we choose a model to run the full analysis with.

```{r choosing the dynamic dataset for analysis}
df_analysis <- df_panel
```

# Standard descriptive statistics

## Sample structure

```{r basic panel structure}
# number of firms, years, observations
df_analysis %>% 
  summarise(N_obs = comma(n()), N_firms = comma(n_distinct(`ISIN Code`)), 
            Years = n_distinct(year), N_countries = comma(n_distinct(`Country of Incorporation`))) %>%
  kable(booktabs = TRUE) %>%
  kable_styling()


# observations per year
df_analysis %>% 
  count(year) %>% 
  mutate(Share = percent(n/sum(n))) %>%
  select(Year=year, Obs=n, Share) %>%
  kable(booktabs=T) %>% kable_styling()
```

```{r N observations by country, treatment status, and GICS sector}
# observations per country
geo_table <- df_analysis %>% 
  group_by(Country = `Country of Incorporation`) %>% 
  summarise(
    Obs = n(),
    `Unique Firms` = n_distinct(`ISIN Code`)
  ) %>% 
  ungroup() %>% 
  arrange(desc(Obs)) %>% 
  mutate(Share = percent(Obs / sum(Obs))) %>%
  select(Country, Obs, `Unique Firms`, Share) %>%
  kable(
    format = "html", 
    booktabs = TRUE, 
    caption = "Geographic Sample Distribution and Composition"
  ) %>% 
  kable_styling(
    bootstrap_options = c("striped", "condensed"), 
    full_width = FALSE,
    font_size = 14
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#444444")

save_kable(geo_table, file = "geographic_distribution.png", zoom = 2)

geo_table


#observations per treatment status
df_analysis %>% 
  count(group) %>% 
  mutate(Share = scales::percent(n / sum(n))) %>% 
  rename(Group = group, Obs = n) %>% 
  kable(booktabs = TRUE) %>% 
  kable_styling(latex_options = "striped") # Optional: adds better visual structure

#observations per GICS sector
df_analysis %>% 
  filter(`GICS Sector Name` != "Financials") %>%
  count(`GICS Sector Name`, sort=T) %>% 
  mutate(Share = percent(n/sum(n))) %>%
  select(Sector=`GICS Sector Name`, Obs=n, Share) %>%
  kable(booktabs=T) %>% kable_styling()

#observations treated per year
df_analysis %>% 
  count(year, treatment_exposed) %>% 
  pivot_wider(names_from = treatment_exposed, values_from = n) %>%
  mutate(Pct_Treated = scales::percent(`TRUE` / (`FALSE` + `TRUE`))) %>%
  select(Year = year, Control = `FALSE`, Treated = `TRUE`, Pct_Treated) %>%
  kable(digits = 0, booktabs = TRUE) %>% 
  kable_styling()

#observation oversight table
#columns: group based on a condition that differs for each group, firm-years, unique firms, countries, treatment year
observation_table <- df_analysis %>%
  group_by(group) %>%
  summarise(
    `Firm-Years` = n(),
    `Unique Firms` = n_distinct(`Company Name`), 
    `Countries` = n_distinct(`Country of Incorporation`),      
    `Treatment Year` = if_else(first(group) == "Treatment", "2022", "N/A"),
    .groups = "drop"
  ) %>%
  arrange(factor(group, levels = c("Treatment", "Control", "Placebo")))

print(observation_table)
```

```{r plot for region}
df_region <- df_panel %>%
  distinct(`Identifier (RIC)`, .keep_all = TRUE) %>%
  mutate(
    region = case_when(
      `Country of Incorporation` %in% c(
        "Austria","Belgium","Cyprus","Denmark","Finland","France","Germany",
        "Greece","Hungary","Ireland","Italy","Luxembourg","Malta",
        "Netherlands","Poland","Portugal","Romania","Slovenia","Spain","Sweden"
      ) ~ "EU",
      
      `Country of Incorporation` %in% c(
        "United Kingdom","Norway","Switzerland","Iceland"
      ) ~ "UK & Europe non-EU",
      
      `Country of Incorporation` %in% c(
        "United States","Canada"
      ) ~ "North America",
      
      `Country of Incorporation` %in% c(
        "China (Mainland)","Hong Kong","India","Indonesia","Japan","Malaysia",
        "Pakistan","Philippines","Singapore","South Korea","Taiwan","Thailand",
        "Vietnam","Sri Lanka","Kazakhstan","New Zealand"
      ) ~ "Asia-Pacific",
      
      `Country of Incorporation` %in% c(
        "Israel","Kuwait","Qatar","Saudi Arabia","United Arab Emirates","Bahrain","Oman","Turkey"
      ) ~ "Middle East",
      
      `Country of Incorporation` %in% c(
        "Nigeria","South Africa","Kenya","Egypt","Morocco"
      ) ~ "Africa",
      
      `Country of Incorporation` %in% c(
        "Argentina","Brazil","Chile","Colombia","Mexico","Peru","Panama"
      ) ~ "Latin America",
      
      `Country of Incorporation` %in% c(
        "Cayman Islands","Jersey","Bermuda","Gibraltar","Guernsey","Faroe Islands",
        "Isle of Man","Marshall Islands","Liberia","British Virgin Islands","Curacao",
        "Mauritius"
      ) ~ "Offshore / other",
      
      is.na(`Country of Incorporation`) ~ "Missing",
      TRUE ~ "Other"
    )
  )

df_region %>%
  count(region) %>%
  mutate(region = reorder(region, n)) %>%
  ggplot(aes(x = region, y = n, fill = region)) +
  geom_col(width = 0.7) +
  coord_flip() +
  labs(
    x = NULL,
    y = "Number of firms",
    title = "Sample composition by region of incorporation"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r by treatment control and placebo firms by region}
df_region %>%
  count(region, group) %>%
  mutate(
    region = fct_reorder(region, n, .fun = sum),
    group = factor(group, levels = c("Control", "Placebo", "Treatment"))
  ) %>%
  ggplot(aes(x = region, y = n, fill = group)) +
  geom_col(width = 0.72, color = "white", linewidth = 0.25) +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "Control" = "#C0392B",
      "Placebo" = "#7F8C8D",
      "Treatment" = "#108A99"
    )
  ) +
  labs(
    x = NULL,
    y = "Number of firms",
    fill = NULL,
    title = "Treatment, control, and placebo firms by region"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 11),
    panel.grid.major.y = element_blank(),
    legend.position = "top",
    legend.justification = "left",
    legend.text = element_text(size = 11)
  )
```

```{r kernal density plot of GW by treatment PRE/POST}
df_density <- df_analysis %>%
  filter() %>%
  mutate(
    period = if_else(year <= 2021, "Pre-treatment (2017–2021)", "Post-treatment (2022–2025)"),
    group = if_else(treatment_exposed == 1, "Treated Group", "Control Group")
  )

ggplot(df_density, aes(x = greenwashing_environmental, fill = period, color = period)) +
  geom_density(alpha = 0.25, linewidth = 1, adjust = 1) +
  facet_wrap(~ group) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "grey50") +
  labs(
    title = "Kernel density of Greenwashing Index before and after CSRD",
    x = "Greenwashing Index",
    y = "Density",
    fill = NULL,
    color = NULL
  ) +
  scale_fill_manual(values = c("Pre-treatment (2017–2021)" = "#1f77b4",
                               "Post-treatment (2022–2025)" = "#C0392B")) +
  scale_color_manual(values = c("Pre-treatment (2017–2021)" = "#1f77b4",
                                "Post-treatment (2022–2025)" = "#C0392B")) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "black") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Descriptive statistics and correlations

```{r descriptive statistics of points of interest}
#define variables and formal names for the table
descr_vars <- c(
  "total_assets"               = "Total Assets (Billion EUR)",
  "roa" = "Return on Assets (net income / total assets)",
  "leverage" = "Financial Leverage (Total debt / Total assets)",
  "share_turnover" = "share turnover / market capitalisation",
  "tobins_q" = "Market Valuation (Market capitalisation / total assets)"
)
descr_vars <- c(setNames("Greenwashing Index", y_var), descr_vars)

#build the table
desc_table <- map_dfr(names(descr_vars), function(v) {
  x <- df_analysis[[v]]
  
  # Scale assets to Billions only for display
  if(v == "total_assets") x <- x / 1e9
  
  data.frame(
    Variable = descr_vars[v],
    N        = sum(!is.na(x)),
    Mean     = mean(x, na.rm = TRUE),
    St.Dev.  = sd(x, na.rm = TRUE),
    Min      = min(x, na.rm = TRUE),
    P25      = quantile(x, 0.25, na.rm = TRUE),
    Median   = median(x, na.rm = TRUE),
    P75      = quantile(x, 0.75, na.rm = TRUE),
    Max      = max(x, na.rm = TRUE)
  )
})

#render the table
desc_table %>%
  kable(format = "html", digits = 3, caption = "Table X: Descriptive Statistics", 
        format.args = list(big.mark = ","),
        col.names = c("Variable", "N", "Mean", "St. Dev.", "Min", "P25", "Median", "P75", "Max")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#444444")
```

## Pre-treatment Parallel trends

```{r distribution of greenwashing pre announcement}
df_descriptive <- df_panel %>%
  filter(year < 2022) %>%     # pre‑treatment
  select(`Identifier (RIC)`, year, group, greenwashing_esg) %>%
  drop_na(greenwashing_esg)

df_descriptive %>%
  mutate(group = fct_relevel(group, "Control", "Placebo", "Treatment")) %>%
  ggplot(aes(x = group, y = greenwashing_esg, fill = group)) +
  geom_boxplot(width = 0.75, linewidth = 0.8, outlier.alpha = 0.3) +
  scale_fill_manual(
    values = c(
      "Control"  = "#C0392B",
      "Placebo"  = "#7F8C8D",
      "Treatment" = "#108A99"
    ),
    guide = "none"
  ) +
  labs(
    title = "Distribution of Greenwashing Index\nby group (pre-announcement period)",
    x = NULL,
    y = "Greenwashing Index (ESG‑based)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.text.x = element_text(size = 12)
  )

df_descriptive %>%
  mutate(group = fct_relevel(group, "Control", "Placebo", "Treatment")) %>%
  ggplot(aes(x = group, y = greenwashing_esg, fill = group)) +
  geom_violin(trim = FALSE, alpha = 0.8) +
  geom_boxplot(
    width = 0.2, outlier.alpha = 0.3,
    fill = "white", linewidth = 0.8
  ) +
  scale_fill_manual(
    values = c(
      "Control"   = "#C0392B",
      "Placebo"   = "#7F8C8D",
      "Treatment" = "#108A99"
    ),
    guide = "none"
  ) +
  labs(
    title = "Distribution of Greenwashing Index\nby group (pre-announcement period)",
    x = NULL,
    y = "Greenwashing Index (ESG‑based)"
  ) 
```

```{r balance test}
# Turn off scientific notation globally for this chunk
options(scipen = 999)

# filter for pre-treatment years
df_pre <- df_analysis %>%
  filter(year >= 2019 & year <= 2021) %>%
  filter(group %in% c("Treatment", "Control"))

table_3_data <- map_dfr(names(descr_vars), function(v) {
  
  # Conduct t-test: variable ~ group
  t_res <- t.test(df_pre[[v]] ~ df_pre$group)
  
  # Extract means (Note: R orders groups alphabetically)
  means <- t_res$estimate
  p_val <- t_res$p.value
  
  # Format p-value cleanly for an academic table
  p_val_formatted <- if(p_val < 0.001) "< 0.001" else as.character(round(p_val, 3))
  
  data.frame(
    Variable = descr_vars[v],
    `Treatment Mean` = round(means[2], 3), 
    `Control Mean` = round(means[1], 3),
    `Diff (T-C)` = round(means[2] - means[1], 3),
    `p-value` = p_val_formatted,
    check.names = FALSE # Prevents R from converting spaces to dots in column names
  )
})

table_3_data %>%
  kable(
    format = "html",
    align = c("l", "r", "r", "r", "r"),
    caption = "Table 3: Pre-Treatment Balance Test (2019–2021 Means)",
    col.names = c("Variable", "Treatment Mean", "Control Mean", "Diff. (T-C)", "p-value")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#444444") %>%
  footnote(general = "p-values are obtained from two-sample Welch t-tests comparing the Treatment and Control groups during the 2019–2021 baseline window.")
```

```{r parallel trends plot}
df_event <- df_analysis %>% 
  filter(group %in% c("Treatment", "Control")) %>%
  # Drop NAs in your variables to prevent the model from crashing
  filter(!is.na(roa), !is.na(leverage), !is.na(total_assets), !is.na(`GICS Sector Code`))

event_study <- feols(
  as.formula(paste0(
    y_var, " ~ i(year, treatment_exposed, ref = 2021)", 
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q | `Identifier (RIC)` + year"
  )),
  data = df_event,
  vcov = ~ `Identifier (RIC)`
)

iplot(
  event_study,
  main = "Event Study: Impact of CSRD on Greenwashing",
  xlab = "Year",
  ylab = "Estimated Effect (Coefficient)"
)
```

Here we observe that the parallel trends hold for each year

```{r parallel trends visualisation}
df_analysis %>%
  filter(year >= 2017) %>%
  mutate(treatment_exposed = as.factor(treatment_exposed)) %>% 
  group_by(year, treatment_exposed) %>%
  summarise(mean_gw = median(.data[[y_var]], na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(year, mean_gw, color = treatment_exposed, group = treatment_exposed)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  geom_vline(xintercept = 2022, linetype = "dashed") +
  geom_vline(xintercept = 2024, linetype = "dashed") +
  annotate("text", x = 2022, y = Inf, label = "CSRD announcement", angle = 90, vjust = 1.2, hjust = 1) +
  annotate("text", x = 2024, y = Inf, label = "CSRD goes in effect", angle = 90, vjust = 1.2, hjust = 1) +
  labs(
    title = "Average Greenwashing temporal trend (2017-2025)", # Note: adjusted title to match filter
    x = "Year",
    y = "Average Greenwashing Index"
  ) +
  theme_minimal()
```

```{r decomposed time trend}
df_trends_long <- df_panel %>%
  filter(year >= 2017) %>%
  group_by(year, treatment_exposed) %>%
  summarise(
    "ESG Disclosure Score" = median(esg_disclosure_z, na.rm = TRUE),
    "ESG Performance Score" = median(esg_performance_z, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Pivot only the two components into long format
  pivot_longer(
    cols = c("ESG Disclosure Score", "ESG Performance Score"),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  mutate(
    group_label = ifelse(treatment_exposed == TRUE, "Treatment", "Control")
  )

ggplot(
  df_trends_long,
  aes(x = year, y = Value, color = group_label, linetype = Metric, group = interaction(group_label, Metric))
) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  
  # Policy lines for CSRD announcement and effect
  geom_vline(xintercept = c(2022, 2024), linetype = "dashed", color = "gray50") +
  
  # Labels for the events
  annotate("text", x = 2022, y = Inf, label = "CSRD announcement", angle = 90, vjust = 1.2, hjust = 1.1, size = 3.5, color = "gray30") +
  annotate("text", x = 2024, y = Inf, label = "CSRD goes in effect", angle = 90, vjust = 1.2, hjust = 1.1, size = 3.5, color = "gray30") +
  
  # Unified color scheme (consistent with paper)
  scale_color_manual(values = c(
    "Control"   = "#C0392B",
    "Treatment" = "#108A99"
  )) +
  
  # Explicit line types to split disclosure and performance cleanly
  scale_linetype_manual(values = c(
    "ESG Disclosure Score"  = "solid",
    "ESG Performance Score" = "dashed"
  )) +
  
  # Titles and labels
  labs(
    title = "ESG Disclosure and Performance Trends over Time",
    subtitle = "Median standardized scores by treatment status",
    x = "Year",
    y = "Median Standardized Score",
    color = NULL,
    linetype = NULL
  ) +
  
  # Minimal, publication‑style theme
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
    panel.grid.minor = element_blank()
  )
```


```{r disclosure vs performance z score}
df_analysis %>%
  filter(year >= 2019) %>%
  group_by(year, treatment_exposed) %>%
  summarise(
    disclosure = mean(esg_disclosure_z, na.rm = TRUE),
    performance = mean(esg_performance_z, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Pivot to long format for clean, consistent plotting and automated legends
  pivot_longer(
    cols = c(disclosure, performance),
    names_to = "component",
    values_to = "value"
  ) %>%
  mutate(
    component = case_when(
      component == "disclosure"  ~ "ESG disclosure score",
      component == "performance" ~ "ESG performance score"
    ),
    # Ensure treatment is treated as a clean factor for faceting labels
    treatment_exposed = case_when(
      treatment_exposed == 1 ~ "Treated", # Adjust these if your dummy labels are different
      treatment_exposed == 0 ~ "Control",
      TRUE ~ as.character(treatment_exposed)
    )
  ) %>%
  ggplot(aes(x = year, y = value, color = component, group = component)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  facet_wrap(~treatment_exposed) +
  # Policy lines consistent with previous plots
  geom_vline(xintercept = 2022, linetype = "dashed", color = "gray50") +
  geom_vline(xintercept = 2024, linetype = "dashed", color = "red") +
  # Event labels
  annotate("text", x = 2022, y = Inf, label = "CSRD announcement", angle = 90, vjust = 1.2, hjust = 1, size = 3) +
  annotate("text", x = 2024, y = Inf, label = "CSRD in effect", angle = 90, vjust = 1.2, hjust = 1, size = 3) +
  # Titles and labels matching your theme
  labs(
    title = "Disclosure Z > Performance Z → Gap Compression",
    x = "Year",
    y = "Mean standardized score (Z-score)",
    color = "Metric"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 11) # Cleans up facet headers
  )
```

# Difference-in-differences (regular type)

## H1 Main effect

```{r base model}
base_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement" #base, only the announcement and treatment exposure
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(base_model)
```

```{r base plus controls}
controls_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q " #include the controls
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(controls_model)
```

```{r base plus controlled plus fixed effects}
fe_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q  |", #controls
    "`Identifier (RIC)` + year" #fixed effects for firm and year
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model)
```

## H2 Mechanism test

```{r the narrative mechanism}
#dependent variable is the disclosure score
mechanism_disclosure <- feols(
  esg_disclosure_z ~ treatment_exposed * post_announcement + log(total_assets) + roa + leverage + share_turnover + tobins_q |
    `Identifier (RIC)` + year,
  data = df_analysis %>% filter(scope1_reported == 1, scope1_emissions_log > 0),
  vcov = ~ `Identifier (RIC)`
)

summary(mechanism_disclosure)
```

```{r the performance mechanism}
#dependent variable is the standardised performance score
mechanism_performance <- feols(
  esg_performance_z ~ treatment_exposed * post_announcement + log(total_assets) + roa + leverage + share_turnover + tobins_q  |
    `Identifier (RIC)` + year,
  data = df_analysis %>% filter(scope1_reported == 1, scope1_emissions_log > 0),
  vcov = ~ `Identifier (RIC)`
)

summary(mechanism_performance)
```


```{r modelled }
estimated_model <- feols(
  scope1_emissions_log ~ treatment_exposed * post_announcement + log(total_assets) + roa + leverage + share_turnover + tobins_q  |
    `Identifier (RIC)` + year,
  data = df_analysis %>% filter(scope1_estimated == 1, scope1_emissions_log > 0),
  vcov = ~ `Identifier (RIC)`
)

summary(estimated_model)
```

```{r mechanism test visualised}
# Calculate means AND count observations to catch the N drop
df_clean_mech <- df_analysis %>%
  filter(year >= 2019, scope1_emissions_log > 0) %>%
  mutate(
    type = case_when(
      scope1_reported == 1 ~ "Actual Reported",
      scope1_estimated == 1 ~ "Third-Party Modelled"
    ),
    treated = ifelse(treatment_exposed == TRUE, "Treated Firms", "Control Firms")
  ) %>%
  filter(!is.na(type)) %>%
  group_by(year, treated, type) %>%
  summarise(
    mean_log = mean(scope1_emissions_log, na.rm = TRUE),
    n_obs = n(),  # Count how many firms are in this cell
    .groups = "drop"
  ) %>%
  # FILTER OUT noisy years where sample size collapses (e.g., fewer than 5 firms)
  filter(n_obs >= 5) 

# Index everything cleanly to 2021 = 0
baselines <- df_clean_mech %>% 
  filter(year == 2021) %>% 
  select(treated, type, base = mean_log)

df_clean_mech %>%
  left_join(baselines, by = c("treated", "type")) %>%
  mutate(indexed = mean_log - base) %>%
  
  # 1. Map color to 'treated' group, and linetype/shape to data 'type'
  ggplot(aes(x = year, y = indexed, color = treated, linetype = type, group = interaction(treated, type))) +
  geom_hline(yintercept = 0, color = "gray60") +
  geom_line(linewidth = 1) +
  geom_point(aes(shape = type), size = 2) + # Distinct point shapes help keep it scannable
  geom_vline(xintercept = c(2022, 2024), linetype = "dashed", color = "gray40") +
  
  # Split columns by Treated vs Control
  facet_wrap(~treated) + 
  
  # 2. Apply your explicit paper color scheme
  # Note: Make sure these names perfectly match your "treated" column strings (e.g., "Treated Firms" vs "Treatment")
  scale_color_manual(values = c(
    "Control Firms" = "#C0392B", 
    "Treated Firms" = "#108A99"
  )) +
  
  # 3. Force solid for Reported and dashed for Modelled
  scale_linetype_manual(values = c(
    "Actual Reported" = "solid", 
    "Third-Party Modelled" = "dashed"
  )) +
  scale_shape_manual(values = c(
    "Actual Reported" = 16, # Solid circle
    "Third-Party Modelled" = 17 # Triangle
  )) +
  
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold") # Cleans up the facet headers
  ) +
  labs(
    title = "Carbon Emissions Trajectory (Indexed to 2021)",
    y = "Change in Log Emissions from 2021 Baseline",
    x = "Year"
  )

df_analysis %>%
  filter(year >= 2019) %>%
  mutate(treated = ifelse(treatment_exposed == TRUE, "Treated", "Control")) %>%
  group_by(year, treated) %>%
  summarise(
    pct_reported = mean(scope1_reported, na.rm = TRUE) * 100,
    .groups = "drop"
  ) %>%
  ggplot(aes(x = year, y = pct_reported, color = treated, group = treated)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  geom_vline(xintercept = c(2022, 2024), linetype = "dashed") +
  labs(
    title = "Mechanism: Shift from Modelled to Actual Reporting",
    y = "% of Firms Disclosing Actual Reported Emissions",
    x = "Year"
  ) +
  theme_minimal()
```


## H3 Industry Heterogeneity

```{r triple d formula}
# Define your formula cleanly
triple_d_formula <- as.formula(paste0(
  y_var, " ~ treatment_exposed * post_announcement * low_emission_sector + ",
  "log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
  "`Identifier (RIC)` + year"
))

# Run the fixed-effects OLS
triple_d <- feols(
  triple_d_formula,
  data = df_analysis,
  vcov = ~ `Identifier (RIC)` # Clusters standard errors at the firm level
)

summary(triple_d)
```

```{r triple d formula with environmental as dep}
# Define your formula cleanly
triple_d_formula <- as.formula(paste0(
  "greenwashing_environmental ~ treatment_exposed * post_announcement * high_emission_sector + ",
  "log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
  "`Identifier (RIC)` + year"
))

# Run the fixed-effects OLS
triple_d <- feols(
  triple_d_formula,
  data = df_analysis,
  vcov = ~ `Identifier (RIC)` # Clusters standard errors at the firm level
)

summary(triple_d)
```



```{r}
#get the list of sectors
sectors <- unique(df_analysis$`GICS Sector Name`)
sectors <- sectors[!is.na(sectors)] #remove NA's 

#run the model for each sector and extract the DiD coefficient
df_gics_quick <- map_dfr(sectors, function(sec) {
  df_sub <- df_analysis %>% filter(`GICS Sector Name` == sec)
  
  tryCatch({
    model <- feols(
      as.formula(paste0(
        y_var, " ~ treatment_exposed * post_announcement + ",
        "log(total_assets) + roa + leverage + share_turnover + tobins_q | `Identifier (RIC)` + year"
      )),
      data = df_sub,
      vcov = ~ `Identifier (RIC)`
    )
    
    ct <- as.data.frame(coeftable(model))
    ci <- as.data.frame(confint(model, level = 0.95))
    row_name <- "treatment_exposedTRUE:post_announcementTRUE"
    
    tibble(
      sector = sec,
      estimate = ct[row_name, 1],
      conf.low = ci[row_name, 1],
      conf.high = ci[row_name, 2]
    )
  }, error = function(e) {
    return(NULL) # Skip if a sector doesn't have enough data to converge
  })
})

#pretty plot
high_emission_macro <- c("Energy", "Materials", "Utilities")

df_gics_plot <- df_gics_quick %>%
  mutate(
    # Check if the sector belongs to the high emission group
    is_high_emission = ifelse(sector %in% high_emission_macro, "High-Emission Sectors", "Other Sectors")
  )

ggplot(df_gics_plot, aes(x = estimate, y = reorder(sector, estimate), color = is_high_emission)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.8) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), size = 0.7, linewidth = 1) +
  
  # Inject manual scale to match your academic paper profile
  # Using Crimson/Red for the high-emission focus group and slate gray for the rest
  scale_color_manual(values = c(
    "High-Emission Sectors" = "#C0392B", 
    "Other Sectors"         = "#7F8C8D"
  )) +
  
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(face = "bold", color = "black"),
    plot.title = element_text(face = "bold", size = 14)
  ) +
  labs(
    title = "CSRD Effect by GICS Sector",
    subtitle = "Point estimates and 95% confidence intervals",
    x = "Estimated Treatment Effect (ATT)",
    y = NULL
  )
```


## Robustness Checks

```{r baseline model for robusntess}
summary(fe_model) #see cell named: "r base plus controlled plus fixed effects"
```

```{r uk firms as control group}
uk_control_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis %>%
    filter(treatment_exposed == TRUE | UK_based == TRUE), #filter the data panel to only include treatment exposed firms, and UK_based firms
  vcov = ~ `Identifier (RIC)`
)

summary(uk_control_model)
```

```{r placebo test with uk as treated group}
#uses the untreated UK group as the treated group to check for spillover/invalid info
df_placebo_uk <- df_analysis %>%
  filter(group %in% c("Placebo", "Treatment")) %>%
  mutate(
    # Ensure this is a factor or numeric for the regression
    fake_treatment = ifelse(group == "Placebo", 1, 0),
    post_val = ifelse(post_announcement == TRUE, 1, 0)
  )

placebo_model <- feols(
  as.formula(paste0(y_var, " ~ fake_treatment * post_val + log(total_assets) + leverage + share_turnover + tobins_q  | `Identifier (RIC)` + year")),
  data = df_placebo_uk,
  vcov = ~ `Identifier (RIC)`
)

summary(placebo_model)
```


```{r placebo treatment T-3}
#this model checks a different treatment year, in this case 2019 or time T - 3
#first we filter the dataset to exclude the treatment period
df_placebo_t3 <- df_analysis %>%
  filter(year < 2022) %>% 
  mutate(fake_post_2019 = ifelse(year >= 2019, 1, 0)) #create fake post period

#then we run the model
placebo_model_2019 <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * fake_post_2019", 
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_placebo_t3,
  vcov = ~ `Identifier (RIC)`
)

summary(placebo_model_2019)
```

```{r no industry z-scoring}
#no clustering by the industry or year, just the raw score
no_cluster_model <- feols(
  as.formula(paste0(
    "greenwashing_esg_raw ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(no_cluster_model)

#year-standardized greenwashing index
y_cluster_model <- feols(
  as.formula(paste0(
    "greenwashing_esg_y ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + share_turnover + tobins_q | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(y_cluster_model)

#industry_standardised greenwashing index
i_cluster_model <- feols(
  as.formula(paste0(
    "greenwashing_esg_i ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + share_turnover + tobins_q | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(i_cluster_model)
```

```{r winsorization }
#managing of extreme outliers, take the bottom and top 5% and set to the 5th percentile value, and 95th percentile respectively 

#function to winsorise
winsorize_5_95 <- function(x) {
  p5 <- quantile(x, 0.05, na.rm = TRUE)
  p95 <- quantile(x, 0.95, na.rm = TRUE)
  pmax(pmin(x, p95), p5)
}

df_analysis <- df_analysis %>%
  mutate(
    !!paste0(y_var, "_wins") := winsorize_5_95(.data[[y_var]]) #winsorised for y_var
  )

robustness_model_wins <- feols(
  as.formula(paste0(
    paste0(y_var, "_wins"), " ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(robustness_model_wins)
```

```{r psm model with fixed effects}
fe_model_psm <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_psm_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_psm)
```


```{r mdm model with fixed effects}
fe_model_mdm <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_mdm_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_mdm)
```

```{r cem model with fixed effects}
fe_model_cem <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_cem_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_cem)
```


```{r summarising table}
robustness_list <- list(
  "Sophisticated model" = fe_model,
  "UK Control" = uk_control_model,
  "Placebo (UK)" = placebo_model,
  "Placebo (T-3)" = placebo_model_2019,
  "Raw GW" = no_cluster_model,
  "Year clustered GW" = y_cluster_model,
  "Industry clustered GW" = i_cluster_model,
  "Winsorized" = robustness_model_wins,
  "Propensity score matching" = fe_model_psm,
  "Coarsed exact matching" = fe_model_cem,
  "MDM matching" = fe_model_mdm
)

cm <- c(
  "treatment_exposedTRUE:post_announcementTRUE" = "DID Interaction (Treat × Post)",
  "fake_treatment:post_val" = "Placebo: UK as Treated",
  "treatment_exposedTRUE:fake_post_2019" = "Placebo: 2019 Treatment"
)

modelsummary(
  robustness_list,
  stars = c('*' = .1, '**' = .05, '***' = .01),
  coef_map = cm,
  gof_omit = 'DF|Deviance|AIC|BIC|Log.Lik|Within|Pseudo|FE',
  title = "Table X: Robustness Analysis of CSRD Impact",
  notes = "Standard errors clustered at firm (RIC) level. Controls included."
)
```


## Forest plot

```{r plot of all coefficients}
# function to extract fixest coefficients cleanly with 95% CIs
extract_fixest <- function(model) {
  ct <- as.data.frame(coeftable(model))
  ci <- as.data.frame(confint(model, level = 0.95))
  
  tibble(
    term = rownames(ct),
    estimate = ct[[1]],      
    conf.low = ci[[1]],      
    conf.high = ci[[2]]      
  )
}

# Combine H1 baseline models with robustness list into one plotting list
master_plot_list <- c(
  list(
    "1. Base Model (No Controls/FE)" = base_model,
    "2. Base Model + Firm Controls"  = controls_model
  ),
  robustness_list
)

# map through the master list
df_plot_data <- map_dfr(master_plot_list, extract_fixest, .id = "specification") %>%
  # filter down to only the primary interaction terms
  filter(term %in% c(
    "treatment_exposedTRUE:post_announcementTRUE",
    "fake_treatment:post_val",
    "treatment_exposedTRUE:fake_post_2019"
  )) %>%
  # clean up labels
  mutate(
    type = case_when(
      specification %in% c("1. Base Model (No Controls/FE)", "2. Base Model + Firm Controls") ~ "H1 Baseline Build-up", #baseline build
      specification == "Sophisticated model" ~ "Main Model (Firm & Year FE)", #main model
      specification %in% c("Placebo (UK)", "Placebo (T-3)") ~ "Placebo Tests", #placebo tests
      specification %in% c("Propensity score matching", "Coarsed exact matching", "MDM matching") ~ "Matching Extensions", #matching algortihtms
      TRUE ~ "Alternative Specifications" #else
    ),
    # clean up names
    specification = case_when(
      specification == "Sophisticated model" ~ "3. Main Model (Firm & Year FE)",
      specification == "Raw GW"              ~ "Alternative DV: Raw GW Index",
      specification == "Year clustered GW"    ~ "Alternative DV: Year Standardized",
      specification == "Industry clustered GW"~ "Alternative DV: Industry Standardized",
      TRUE ~ specification
    )
  )

# generate the forest plot
ggplot(df_plot_data, aes(x = estimate, y = fct_rev(fct_inorder(specification)), color = type)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.8) + #vertical line at 0
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), size = 0.7, linewidth = 1) + #add point estimates and 95% intervals
  scale_color_manual(values = c(
    "H1 Baseline Build-up" = "#34495E",      
    "Main Model (Firm & Year FE)" = "#1F77B4",
    "Alternative Specifications" = "#2CA02C",
    "Matching Extensions" = "#9467BD",   
    "Placebo Tests" = "#D62728"
  )) +
  labs(
    title = "Stability of the CSRD Treatment Effect (ATT)",
    subtitle = "Point estimates and 95% confidence intervals across baseline specifications, matching models, and placebos",
    x = "Estimated Treatment Coefficient (ATT Effect Size)",
    y = NULL,
  ) +
  
  #adjustments
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(), # Removes horizontal lines for a cleaner profile
    axis.text.y = element_text(face = "bold", color = "black"),
    legend.title = element_blank(),
    legend.position = "bottom"
  )
```

