---
title: "Analysis"
author: "Melle Stephen Oudshoorn"
date: "`r Sys.Date()`"
output: html_document
---

# Setup and work

```{r setup chunk, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)
```


```{r load packages and key lists}
library(tidyverse)
library(ggplot2)
library(stringr)
library(modelsummary)
library(psych)
library(patchwork)
library(fixest)
library(MatchIt)
#pretty tables
library(knitr)
library(kableExtra)
library(scales)
library(car)
library(gt)

eu_countries <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czech Republic",
  "Denmark", "Estonia", "Finland", "France", "Germany", "Greece", "Hungary",
  "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Malta", "Netherlands",
  "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", "Spain", "Sweden"
)

high_emission_sectors <- c( #division of high emissions sectors
      "Oil, Gas & Consumable Fuels",
      "Energy Equipment & Services",
      "Metals & Mining", 
      "Chemicals",
      "Paper & Forest Products",
      "Construction Materials",
      "Electric Utilities",
      "Gas Utilities",
      "Water Utilities",
      "Multi-Utilities",
      "Independent Power and Renewable Electricity Producers"
    )
```

```{r setup the basics}
df_panel <- read.csv("data/CSRD_thesis_data/panel_final.csv", check.names = FALSE) %>% 
  filter(!is.na(esg_score)) %>%  #filter out NA's for the ESG score
  filter(`GICS Sector Name` != "Financials") %>% #filter out the financial sector due to unreliable emissions
  filter(year >= 2014, year <= 2025) %>% #2026 and 2013 were in the original dataset but too sparse (around 100 observations each that align with the restrictives)
  filter(!is.na(total_debt_actual)) %>% 
  mutate(
    leverage = total_debt_actual/market_capitalization, #create leverage variable
    roa = net_income_after_tax / total_assets, #create roa variable
    high_emission_sector = `GICS Industry Name` %in% high_emission_sectors, #defines high emission sectors
    is_listed_eu = `Country of Exchange` %in% eu_countries, #defines whether a company is listed in the eu
    #size standards
    has_high_employees = ifelse(is.na(number_of_employees), FALSE, number_of_employees > 500), #whether a company exceeds NFRD size threshold regarding employees
    has_high_assets = ifelse(is.na(total_assets), FALSE, total_assets > 20000000), #whether a company exceeds NFRD size threshold regarding assets
    has_high_turnover = ifelse(is.na(turnover), FALSE, turnover > 40000000), #whether a company exceeds NFRD size threshold regarding turnover
    nfrd_size = has_high_employees, #if they have larger than 500 employees they are exposed
    #eu domicile
    in_EU_EU_domicile = `Country of Incorporation` %in% eu_countries,
    UK_based = `Country of Incorporation` %in% "United Kingdom",
    #NFRD/CSRD exposed
    treatment_exposed = has_high_employees & (has_high_assets | has_high_turnover) &
  (is_listed_eu | in_EU_EU_domicile),  #is in the EU
    #policy shock
    post_csrd = year >= 2024,
    post_announcement = year >= 2022,
    placebo_post_2019 = year >= 2019,
    #emissions
    total_emissions_log = log(scope_1_and_scope_2_and_scope_3_combined_reported_or_estimated_total + 1),
    scope1_emissions_log = log(scope_1_combined_reported_or_estimated_total + 1),
    scope2_emissions_log = log(scope_2_combined_reported_or_estimated_total + 1),
    scope3_emissions_log = log(scope_3_combined_reported_or_estimated_total + 1),
    inv_emissions = -total_emissions_log, #inverted so that higher = better performance
    #scope1 reported or estimated
    scope1_reported = if_else(scope_1_combined_reported_or_estimated_method == "Reported_value", 1L, 0L,
      missing = NA_integer_),
    scope1_estimated = if_else(scope_1_combined_reported_or_estimated_method %in% c(
    "Aggregated_model","Winsorized_low","Extrapolated","Winsorized_high","Energy_extrapolated","Energy_model"),
    1L, 0L, missing = NA_integer_),
    #other pollutants
    nox_log = log(nox_emissions + 1),
    sox_log = log(sox_emissions + 1),
    waste_log = log(hazardous_waste + 1),
    #standardise emissions to match
    total_emissions_std = as.numeric(scale(total_emissions_log)),
    scope1_emissions_std = as.numeric(scale(scope1_emissions_log)),
    scope2_emissions_std = as.numeric(scale(scope2_emissions_log)),
    scope3_emissions_std = as.numeric(scale(scope3_emissions_log)),
    nox_std = as.numeric(scale(nox_log)),
    sox_std = as.numeric(scale(sox_log)),
    waste_std = as.numeric(scale(waste_log)),
    esg_std = as.numeric(scale(esg_score)),
    env_pillar_std = as.numeric(scale(environmental_pillar_score)),
    #group
    group = case_when(
      #treatment
      treatment_exposed == TRUE ~ "Treatment",
      #placebo
      UK_based == TRUE ~ "Placebo",
      #control (excl. UK)
      treatment_exposed == FALSE & UK_based == FALSE ~ "Control", # and UK_based == FALSE
      #fallback
      TRUE ~ "Other"
    )
  )

df_panel <- df_panel %>%
  group_by(`Identifier (RIC)`) %>%
  mutate(
    # Average % change in log emissions only for the pre-treatment years (2019-2021)
    emissions_trajectory = mean(
      ((total_emissions_log - lag(total_emissions_log)) / lag(total_emissions_log))[year %in% 2019:2021], 
      na.rm = TRUE
    )
  ) %>%
  # Fill NAs with 0 to prevent row-loss in regressions
  mutate(emissions_trajectory = replace_na(emissions_trajectory, 0)) %>%
  ungroup()

environmental_vars <- c(
  "scope_1_combined_reported_or_estimated_total",
  "scope_2_combined_reported_or_estimated_total",
  "scope_3_combined_reported_or_estimated_total",
  "nox_emissions",
  "sox_emissions",
  "hazardous_waste",
  "energy_use_total",
  "renewable_energy_use_ratio",
  "waste_recycled_to_total_waste",
  "water_discharged",
  "water_withdrawal_total",
  "environmental_restoration_initiatives",
  "biodiversity_due_diligence",
  "policy_waste",
  "water_management_plan_details",
  "renewable_energy_use",
  "policy_resource_efficiency",
  "climate_policy_statement"
)

social_vars <- c(
  "number_of_employees",
  "women_employees",
  "women_managers",
  "training_hours_total",
  "health_safety_policy",
  "salary_gap",
  "gender_pay_gap_percentage",
  "human_rights_policy",
  "csr_sustainability_external_audit",
  "supply_chain_health_safety_improvements",
  "donations_total",
  "policy_board_diversity"
)

governance_vars <- c(
  "shareholders_score",
  "shareholder_approval_significant_transactions",
  "independent_board_members",
  "csr_sustainability_reporting",
  "csr_sustainability_committee",
  "corruption_oversight",
  "stakeholder_engagement",
  "audit_committee_independence",
  "policy_executive_compensation_esg_performance",
  "whistleblower_protection",
  "policy_bribery_and_corruption"
)

all_vars <- c(environmental_vars, social_vars, governance_vars)
```

```{r build the disclosure score}
construct_disclosure_scores <- function(df, environmental_vars, social_vars, governance_vars) {
  
  score_var <- function(x) {
    is_binary <- is.logical(x) || all(x %in% c(TRUE, FALSE, 1, 0, NA))
    if (is_binary) return(as.integer(!is.na(x) & as.logical(x)))
    return(as.integer(!is.na(x)))
  }

  # Score the matrices
  env_scores <- df[environmental_vars] %>% mutate(across(everything(), score_var))
  soc_scores <- df[social_vars]        %>% mutate(across(everything(), score_var))
  gov_scores <- df[governance_vars]    %>% mutate(across(everything(), score_var))

  # THE FIX: Treat Scope 1, 2, 3 as ONE POINT total.
  # This removes the artificial "3-point floor" and the "0-weighted spike."
  scope_vars <- c("scope_1_combined_reported_or_estimated_total",
                  "scope_2_combined_reported_or_estimated_total",
                  "scope_3_combined_reported_or_estimated_total")
  
  # Calculate Environmental: (Everything else) + (1 point if ANY scope is reported)
  env_sum <- rowSums(env_scores %>% select(-all_of(scope_vars))) + 
             as.integer(rowSums(env_scores[scope_vars]) > 0)

  df <- df %>%
    mutate(
      environmental_disclosure_score = env_sum,
      social_disclosure_score        = rowSums(soc_scores),
      governance_disclosure_score    = rowSums(gov_scores),
      # Total score is now a clean integer scale
      esg_disclosure_score           = environmental_disclosure_score +
                                       social_disclosure_score +
                                       governance_disclosure_score
    )
  
  return(df)
}

# run the calculation
df_panel <- construct_disclosure_scores(
  df_panel,
  environmental_vars,
  social_vars,
  governance_vars
)
```

```{r describe the disclosure variables}
psych::describe(df_panel %>%
  select(
    esg_disclosure_score,
    environmental_disclosure_score,
    social_disclosure_score,
    governance_disclosure_score
  ))

df_esg_disclosure_long <- df_panel %>%
  select(
    esg_disclosure_score,
    environmental_disclosure_score,
    social_disclosure_score,
    governance_disclosure_score
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )

ggplot(df_esg_disclosure_long, aes(x = value)) +
  geom_histogram(bins = 50) +
  facet_wrap(~ variable, scales = "free") +
  labs(
    title = "Distribution of ESG Variables",
    x = NULL,
    y = "Count"
  ) +
  theme_minimal()
```

```{r computing the greenwashing index according to yu et al.}
df_panel <- df_panel %>%
  group_by(year, `GICS Sector Code`) %>%  #group by year and sector to remove mechanical industry variation
  mutate(
    #standardise the disclosure scores within industry-year groups
    esg_disclosure_z = as.numeric(scale(esg_disclosure_score)),
    environmental_disclosure_z = as.numeric(scale(environmental_disclosure_score)),
    social_disclosure_z = as.numeric(scale(social_disclosure_score)),
    governance_disclosure_z = as.numeric(scale(governance_disclosure_score)),
    #standardise performance scores
    esg_performance_z = as.numeric(scale(esg_score)),
    environmental_performance_z = as.numeric(scale(environmental_pillar_score)),
    social_performance_z = as.numeric(scale(social_pillar_score)),
    governance_performance_z = as.numeric(scale(governance_pillar_score)),
    
    #compute greenwashing indeces (standardised disclosure - standardised performance)
    
    #baseline greenwashing index
    greenwashing_esg = esg_disclosure_z - esg_performance_z,
    
    #pillared greenwashing index
    greenwashing_environmental = environmental_disclosure_z - environmental_performance_z,
    greenwashing_social = social_disclosure_z - social_performance_z,
    greenwashing_governance = governance_disclosure_z - governance_performance_z,
    
    #robustness check
    greenwashing_environmental_alt = esg_disclosure_z - environmental_performance_z #robustness check
  ) %>% 
  ungroup()

#full sample alternative (robustness)
df_panel <- df_panel %>%
  mutate(
    esg_disclosure_z_raw = as.numeric(scale(esg_disclosure_score)),
    esg_performance_z_raw = as.numeric(scale(esg_score)),
    greenwashing_esg_raw = esg_disclosure_z_raw - esg_performance_z_raw
  )

#year clustered alternative (robustness)
df_panel <- df_panel %>%
  group_by(year) %>%
  mutate(
    esg_disclosure_z_y = as.numeric(scale(esg_disclosure_score)),
    esg_performance_z_y = as.numeric(scale(esg_score)),
    greenwashing_esg_y = esg_disclosure_z_y - esg_performance_z_y
  ) %>%
  ungroup()
```

```{r computing the greenwashing index according to Dorfleitner & Utz}
#esg vs total ghg
gw_model <- feols(esg_std ~ total_emissions_std | `Identifier (RIC)` + year,
                  cluster = ~`Identifier (RIC)` + year,
                  data = df_panel)
#E-pillar of esg vs total ghg
gw_env_model <-  feols(environmental_pillar_score ~ total_emissions_std | `Identifier (RIC)` + year,
                       cluster = ~`Identifier (RIC)` + year,
                       data = df_panel)

#extracting the residuals safely

#compute the greenwashing as the residual
df_panel <- df_panel %>%
  mutate(
    # ESG-based residual GW
    gw_dorfleitner_utz = esg_std - predict(gw_model, newdata = .),
    
    # Environment-only residual GW
    gw_env_dorfleitner_utz = env_pillar_std - predict(gw_env_model, newdata = .)
  )
```

# Standard descriptive statistics

## Sample structure

```{r basic panel structure}
# number of firms, years, observations
df_panel %>% 
  summarise(N_obs = comma(n()), N_firms = comma(n_distinct(`ISIN Code`)), 
            Years = n_distinct(year), N_countries = comma(n_distinct(`Country of Incorporation`))) %>%
  kable(booktabs = TRUE) %>%
  kable_styling()


# observations per year
df_panel %>% 
  count(year) %>% 
  mutate(Share = percent(n/sum(n))) %>%
  select(Year=year, Obs=n, Share) %>%
  kable(booktabs=T) %>% kable_styling()
```

```{r N observations by country, treatment status, and GICS sector}
# observations per country
df_panel %>% 
  count(`Country of Incorporation`, sort=T) %>% 
  head(10) %>% 
  mutate(Share = percent(n/sum(n))) %>%
  select(Country=`Country of Incorporation`, Obs=n, Share) %>%
  kable(booktabs=T) %>% kable_styling()

#observations per treatment status
df_panel %>% 
  count(group) %>% 
  mutate(Share = scales::percent(n / sum(n))) %>% 
  rename(Group = group, Obs = n) %>% 
  kable(booktabs = TRUE) %>% 
  kable_styling(latex_options = "striped") # Optional: adds better visual structure

#observations per GICS sector
df_panel %>% 
  filter(`GICS Sector Name` != "Financials") %>%
  count(`GICS Sector Name`, sort=T) %>% 
  mutate(Share = percent(n/sum(n))) %>%
  select(Sector=`GICS Sector Name`, Obs=n, Share) %>%
  kable(booktabs=T) %>% kable_styling()

#observations treated per year
df_panel %>% 
  count(year, treatment_exposed) %>% 
  pivot_wider(names_from = treatment_exposed, values_from = n) %>%
  mutate(Pct_Treated = scales::percent(`TRUE` / (`FALSE` + `TRUE`))) %>%
  select(Year = year, Control = `FALSE`, Treated = `TRUE`, Pct_Treated) %>%
  kable(digits = 0, booktabs = TRUE) %>% 
  kable_styling()

#observation oversight table
#columns: group based on a condition that differs for each group, firm-years, unique firms, countries, treatment year
observation_table <- df_panel %>%
  group_by(group) %>%
  summarise(
    `Firm-Years` = n(),
    `Unique Firms` = n_distinct(`Company Name`), 
    `Countries` = n_distinct(`Country of Incorporation`),      
    `Treatment Year` = if_else(first(group) == "Treatment", "2022", "N/A"),
    .groups = "drop"
  ) %>%
  arrange(factor(group, levels = c("Treatment", "Control", "Placebo")))

print(observation_table)
```

```{r visualisations of sample}
# 1. Stacked bar: Year × Treatment
p1 <- df_panel %>%
  count(year, treatment_exposed) %>%
  mutate(group = ifelse(treatment_exposed, "Treated", "Control")) %>%
  ggplot(aes(year, n, fill = group)) +
  geom_col(position = "stack") +
  geom_vline(xintercept = 2024, linetype = "dashed", color = "red") +
  labs(title = "Panel Structure by Treatment", x = "Year", y = "Observations") +
  theme_minimal() +
  theme(legend.position = "bottom")

# 2. EU vs Non-EU bar
p2 <- df_panel %>%
  mutate(eu = `Country of Incorporation` %in% eu_countries) %>%
  count(eu) %>%
  mutate(group = ifelse(eu, "EU", "Non-EU")) %>%
  ggplot(aes(reorder(group, n), n, fill = group)) +
  geom_col() +
  labs(title = "EU vs Non-EU Firms", x = "", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")

# 3. Sector by treatment (dodge)
p3 <- df_panel %>%
  filter(`GICS Sector Name` != "Financials") %>%
  count(`GICS Sector Name`, treatment_exposed) %>%
  ggplot(aes(reorder(`GICS Sector Name`, n), n, fill = treatment_exposed)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(title = "Sectors by Treatment", x = "", y = "Count") +
  theme_minimal()

# Arrange 1 row
p1 | p2 | p3
```

## Greenwashing index

```{r list of greenwashing metrics}
greenwashing_metrics <- c("greenwashing_esg", "greenwashing_environmental", "greenwashing_environmental_alt", "greenwashing_social", "greenwashing_governance", "gw_dorfleitner_utz", "gw_env_dorfleitner_utz")
```


```{r summary stats}
df_panel %>%
  select(all_of(greenwashing_metrics)) %>%
  gather(key = "Measure", value = "GW") %>%
  group_by(Measure) %>%
  summarise(
    Mean = round(mean(GW, na.rm=T), 3),
    SD = round(sd(GW, na.rm=T), 3),
    Median = round(median(GW, na.rm=T), 3),
    P10 = round(quantile(GW, 0.1, na.rm=T), 3),
    P90 = round(quantile(GW, 0.9, na.rm=T), 3),
    .groups = "drop"
  ) %>%
  kable(booktabs=T) %>% kable_styling()
```


### Choosing which metric to use

```{r correlation matrix of greenwashing metrics}
df_panel %>%
  select(all_of(greenwashing_metrics)) %>%
  cor(use = "complete.obs") %>%
  round(3) %>%
  kable(booktabs=T, caption = "Greenwashing Measures Correlation", digits=3) %>%
  kable_styling()
```

```{r kernal density plot of GW by treatment PRE/POST}
df_density <- df_panel %>%
  filter(year >= 2019) %>%
  mutate(
    period = if_else(year <= 2021, "Pre-treatment (2019–2021)", "Post-treatment (2022–2025)"),
    group = if_else(treatment_exposed == 1, "Treated Group", "Control Group")
  )

ggplot(df_density, aes(x = greenwashing_environmental, fill = period, color = period)) +
  geom_density(alpha = 0.25, linewidth = 1, adjust = 1) +
  facet_wrap(~ group) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "grey50") +
  labs(
    title = "Kernel density of Greenwashing Index before and after CSRD",
    x = "Greenwashing Index",
    y = "Density",
    fill = NULL,
    color = NULL
  ) +
  scale_fill_manual(values = c("Pre-treatment (2019–2021)" = "#1f77b4",
                               "Post-treatment (2022–2025)" = "#d62728")) +
  scale_color_manual(values = c("Pre-treatment (2019–2021)" = "#1f77b4",
                                "Post-treatment (2022–2025)" = "#d62728")) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "black") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r metrics correlation with fines}
fines_corrs <- sapply(greenwashing_metrics, function(metric) {
  cor(df_panel$greenwashing_fines_controversies_count, 
      df_panel[[metric]], use = "complete.obs")
})

did_r2 <- numeric(length(greenwashing_metrics))
for(i in seq_along(greenwashing_metrics)) {
  metric <- greenwashing_metrics[i]
  formula_str <- paste0("roa ~ treatment_exposed * post_csrd + ", metric, " | `Identifier (RIC)` + year")
  model <- feols(as.formula(formula_str), data = df_panel, cluster = ~`Identifier (RIC)`)
  did_r2[i] <- r2(model)
}

data.frame(
  Measure = greenwashing_metrics,
  Corr_Fines = round(fines_corrs, 3),
  DiD_R2 = round(did_r2, 4)
) %>%
  arrange(desc(Corr_Fines)) %>%
  kable(booktabs = TRUE, digits = 3, caption = "GW Measure Validation") %>%
  kable_styling()
```

### Descriptives of the chosen metric

```{r choice of the metric}
y_var <- "greenwashing_esg"
```


```{r gw-environmental-descriptives, results='asis'}
df_panel %>%
  summarise(
    `Mean` = round(mean(.data[[y_var]], na.rm = TRUE), 3),
    `SD` = round(sd(.data[[y_var]], na.rm = TRUE), 3),
    `P10` = round(quantile(.data[[y_var]], 0.1, na.rm = TRUE), 3),
    `P50` = round(median(.data[[y_var]], na.rm = TRUE), 3),
    `P90` = round(quantile(.data[[y_var]], 0.9, na.rm = TRUE), 3),
    `N` = comma(sum(!is.na(.data[[y_var]])))
  ) %>%
  kable(booktabs = TRUE, caption = "**Table 2: Primary Outcome - Environmental Greenwashing**") %>%
  kable_styling()
```

## Descriptive statistics and correlations

```{r descriptive statistics of points of interest}
#define variables and formal names for the table
descr_vars <- c(
  "greenwashing_esg" = "Greenwashing Index",
  "total_assets"               = "Total Assets (EUR)",
  "roa" = "Return on Assets (net income / total assets)",
  "leverage" = "Financial Leverage (Total debt / Total assets)",
  "emissions_trajectory" = "Pre-treatment carbon trend (2019-2021)"
)

#build the table
desc_table <- map_dfr(names(descr_vars), function(v) {
  x <- df_panel[[v]]
  
  # Scale assets to Billions only for display
  if(v == "total_assets") x <- x / 1e9
  
  data.frame(
    Variable = descr_vars[v],
    N        = sum(!is.na(x)),
    Mean     = mean(x, na.rm = TRUE),
    St.Dev.  = sd(x, na.rm = TRUE),
    Min      = min(x, na.rm = TRUE),
    P25      = quantile(x, 0.25, na.rm = TRUE),
    Median   = median(x, na.rm = TRUE),
    P75      = quantile(x, 0.75, na.rm = TRUE),
    Max      = max(x, na.rm = TRUE)
  )
})

#render the table
desc_table %>%
  kable(format = "html", digits = 3, caption = "Table X: Descriptive Statistics", 
        format.args = list(big.mark = ","),
        col.names = c("Variable", "N", "Mean", "St. Dev.", "Min", "P25", "Median", "P75", "Max")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#444444")
```

## Pre-treatment Parallel trends

```{r balance test}
#filter for pre-treatment years
df_pre <- df_panel %>%
  filter(year >= 2019 & year <= 2021) %>%
  filter(group %in% c("Treatment", "Control"))

table_3_data <- map_dfr(names(descr_vars), function(v) {
  
  # Conduct t-test: variable ~ group
  t_res <- t.test(df_pre[[v]] ~ df_pre$group)
  
  # Extract means (Note: R orders groups alphabetically, so check order)
  # Usually: [1] is Control, [2] is Treatment if alphabetically sorted
  means <- t_res$estimate
  p_val <- t_res$p.value
  
  data.frame(
    Variable = descr_vars[v],
    `Treatment Mean` = means[2], # Adjust index based on your group names
    `Control Mean` = means[1],
    `Diff (T-C)` = means[2] - means[1],
    `p-value` = p_val
  )
})

table_3_data %>%
  kable(
    format = "html",
    digits = 3,
    caption = "Table 3: Pre-Treatment Balance Test (2019–2021 Means)",
    col.names = c("Variable", "Treatment Mean", "Control Mean", "Diff. (T-C)", "p-value")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#444444") %>%
  # Add a footnote to explain the p-values
  footnote(general = "p-values are from two-sample t-tests comparing Treatment and Control groups during the 2019–2021 period.")
```


```{r}
df_event <- df_panel %>% 
  filter(group %in% c("Treatment", "Control"))

event_study <- feols(
  as.formula(paste0(
    y_var, " ~ i(year, treatment_exposed, ref = 2022) + ",
    "log(total_assets) + leverage + roa | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_event,
  vcov = ~ `Identifier (RIC)`
)

iplot(
  event_study,
  main = "Event Study: Impact of CSRD on Greenwashing",
  xlab = "Year",
  ylab = "Estimated Effect (Coefficient)"
)
```



```{r parallel trends test}
print("Parallel trends SD (pre-2022):")

df_panel %>% #panel
  filter(year < 2022) %>% #filter pre-treatment
  group_by(year, treatment_exposed) %>% #group by year and treatment exposed
  summarise(mean_gw = mean(.data[[y_var]], na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = treatment_exposed, values_from = mean_gw) %>%
  mutate(diff = `TRUE` - `FALSE`) %>%
  summarise(sd_diff = sd(diff, na.rm = TRUE), n_years = n()) %>%
  mutate(t_stat = 0 / sd_diff * sqrt(n_years))
```

```{r parallel trends visualisation}
df_panel %>%
  filter(year >= 2019) %>%
  group_by(year, treatment_exposed) %>%
  summarise(mean_gw = median(.data[[y_var]], na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(year, mean_gw, color = treatment_exposed)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = 2022, linetype = "dashed") +
  geom_vline(xintercept = 2024, linetype = "dashed") +
  annotate("text", x = 2022, y = Inf, label = "CSRD announcement", angle = 90, vjust = 1.2, hjust = 1) +
  annotate("text", x = 2024, y = Inf, label = "CSRD goes in effect", angle = 90, vjust = 1.2, hjust = 1) +
  labs(
    title = "Average Greenwashing temporal trend (2019-2025)",
    x = "Year",
    y = "Average Greenwashing Index"
  ) +
  theme_minimal()
```

# ESG Disclosure and performance scores

```{r environmental scores by year and treatment}
df_panel %>%
  mutate(period = ifelse(year < 2024, "Pre-CSRD", "Post-CSRD")) %>%
  group_by(period, treatment_exposed) %>%
  summarise(
    `Disclosure Z` = round(mean(esg_disclosure_z, na.rm=T), 3),
    `Performance Z` = round(mean(esg_performance_z, na.rm=T), 3),
    `Env Disclosure Z` = round(mean(environmental_disclosure_z, na.rm=T), 3),
    `Env Perf Z` = round(mean(environmental_performance_z, na.rm=T), 3),
    Greenwashing = round(mean(greenwashing_environmental, na.rm=T), 3),
    .groups="drop"
  ) %>%
  kable(caption="**Table 2: Normalized Scores (Z-scores)**") %>%
  kable_styling()
```

```{r disclosure vs performance z score}
df_panel %>%
  filter(year >= 2019) %>%
  group_by(year, treatment_exposed) %>%
  summarise(
    disc_z = mean(environmental_disclosure_z, na.rm=T),
    perf_z = mean(environmental_performance_z, na.rm=T),
    .groups="drop"
  ) %>%
  ggplot(aes(year)) +
  geom_line(aes(y=disc_z, color="Disclosure Z"), size=1) +
  geom_line(aes(y=perf_z, color="Performance Z"), size=1) +
  facet_wrap(~treatment_exposed) +
  geom_vline(xintercept=2024, linetype="dashed", color="red") +
  labs(y="Z-score", title="Disclosure Z > Performance Z → Gap Compression") +
  theme_minimal()
```

```{r country distribution of ESG disclosure vs ESG performance scores}
df_panel %>%
  filter(year >= 2019) %>%
  ggplot(aes(x = environmental_performance_z,
             y = environmental_disclosure_z)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    x = "Environmental Performance (Z-score)",
    y = "Environmental Disclosure (Z-score)",
    title = "Disclosure vs Performance (Z-scores)"
  ) +
  theme_minimal()
```

```{r}
df_panel %>%
  filter(year <= 2023) %>%
  group_by(`Country of Headquarters`) %>%
  summarise(
    disc_z = mean(environmental_disclosure_score, na.rm = TRUE),
    perf_z = mean(environmental_pillar_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = disc_z, y = perf_z)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    x = "Environmental Disclosure",
    y = "Environmental Performance",
    title = "Country-Level Disclosure vs Performance (pre-2024)"
  ) +
  geom_text(aes(label = `Country of Headquarters`), hjust = 1, vjust = 1, size = 3) +
  theme_minimal()
```

```{r boxplots of ESG performance and ESG disclosure score for sample firms}
df_panel %>%
  select(environmental_pillar_score, environmental_disclosure_score) %>%
  pivot_longer(cols = everything(),
               names_to = "metric",
               values_to = "value") %>%
  ggplot(aes(x = metric, y = value, fill = metric)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_x_discrete(labels = c(
    environmental_pillar_score = "Environmental Performance",
    environmental_disclosure_score = "Environmental Disclosure"
  )) +
  labs(
    x = "",
    y = "Score",
    title = "Distribution of ESG Performance vs Disclosure (Full Sample)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r distribution of relative greenwashing scores of sample firms}
df_panel %>%
  ggplot(aes(x = .data[[y_var]])) +
  geom_histogram(bins = 40, alpha = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    x = "Relative Greenwashing Score (Environmental Disclosure Z score - Environmental Performance Z score)",
    y = "Count",
    title = "Distribution of Relative Greenwashing Scores (Full Sample)"
  ) +
  theme_minimal()
```

## Emissions analysis

```{r emissions summary table}
df_panel %>%
  mutate(period = ifelse(year < 2024, "Pre-CSRD", "Post-CSRD")) %>%
  group_by(period, treatment_exposed) %>%
  summarise(
    `Total Log` = round(mean(total_emissions_log, na.rm=T), 3),
    `Scope1 Log` = round(mean(scope1_emissions_log, na.rm=T), 3),
    `Scope2 Log` = round(mean(scope2_emissions_log, na.rm=T), 3),
    `Scope3 Log` = round(mean(scope3_emissions_log, na.rm=T), 3),
    .groups="drop"
  ) %>%
  kable(caption="**Table 4: Log Emissions Pre/Post CSRD**  
  *H2 Expectation: Flat post-CSRD (no real effects)*") %>%
  kable_styling()
```

```{r emissions time series}
df_panel %>%
  filter(year >= 2018) %>%
  group_by(year, treatment_exposed) %>%
  summarise(
    scope1 = mean(scope1_emissions_log, na.rm=T),
    scope2 = mean(scope2_emissions_log, na.rm=T),
    total = mean(total_emissions_log, na.rm=T),
    .groups="drop"
  ) %>%
  pivot_longer(c(scope1, scope2, total), names_to="scope", values_to="log_emissions") %>%
  ggplot(aes(year, log_emissions, color=treatment_exposed)) +
  geom_line() +
  geom_vline(xintercept=2024, linetype="dashed") +
  facet_wrap(~scope) +
  labs(y="Log Emissions", title="Figure 4: Emissions Flat Post-CSRD 
  Score compression, not real effects") +
  theme_minimal()
```

```{r reporting method mix}
df_panel %>%
  mutate(reported_scope1 = case_when(
    scope_1_combined_reported_or_estimated_method == "Reported_value" ~ "Reported",
    TRUE ~ "Estimated/Modelled"
  )) %>%
  group_by(year, treatment_exposed, reported_scope1) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(year, treatment_exposed) %>%
  mutate(share = n / sum(n)) %>%
  ggplot(aes(year, share, fill = reported_scope1)) +
  geom_col() +
  facet_wrap(~treatment_exposed) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Share of Firms (%)", title = "Scope 1: Reported vs Modelled") +
  theme_minimal()
```

## Industry Heterogeneity

```{r industry heterogeneity summary}
df_panel %>%
  filter(treatment_exposed == TRUE) %>%
  mutate(period = ifelse(year < 2024, "Pre-CSRD", "Post-CSRD")) %>%
  group_by(`GICS Sector Name`, period) %>%
  summarise(
    mean_gw = mean(.data[[y_var]], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_gw)) %>%
  mutate(mean_gw = round(mean_gw, 3)) %>%
  gt()
```

```{r bar chart sector motivation}
df_panel %>%
  filter(treatment_exposed == TRUE) %>%
  mutate(period = ifelse(year < 2024, "Pre-CSRD", "Post-CSRD")) %>%
  group_by(`GICS Sector Name`, period) %>%
  summarise(mean_gw = mean(.data[[y_var]], na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = `GICS Sector Name`, y = mean_gw, fill = period)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(
    x = "GICS Sector",
    y = "Mean Greenwashing Index",
    title = "Greenwashing by Sector: Treated Firms Only"
  ) +
  theme_minimal()
```

```{r table sector motivation}
table_industry_gw <- df_panel %>%
  group_by(`GICS Sector Name`) %>%
  summarise(
    n_firms = n_distinct(`Identifier (RIC)`),
    greenwashing_score = mean(.data[[y_var]], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    perc_of_all_firms = 100 * n_firms / sum(n_firms)
  ) %>%
  select(
    Sector = `GICS Sector Name`,
    perc_of_all_firms,
    greenwashing_score
  ) %>%
  arrange(desc(perc_of_all_firms)) %>%
  gt() %>%
  fmt_number(
    columns = c(perc_of_all_firms, greenwashing_score),
    decimals = 3
  ) %>%
  cols_label(
    Sector = "Sector",
    perc_of_all_firms = "Share of Firms (%)",
    greenwashing_score = paste("Average", y_var)
  ) %>%
  tab_header(
    title = "Sector Share of Firms and Average Greenwashing Score"
  )

table_industry_gw
```


# Difference-in-differences (regular type)

## H1 Main effect

```{r base model}
base_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement" #base, only the announcement and treatment exposure
  )),
  data = df_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(base_model)
```

```{r base plus controls}
controls_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + emissions_trajectory" #include the controls
  )),
  data = df_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(controls_model)
```

```{r base plus controlled plus fixed effects}
fe_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + emissions_trajectory |", #controls
    "`Identifier (RIC)` + year" #fixed effects for firm and year
  )),
  data = df_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model)
```

## H2 Mechanism test

```{r reported model}
reported_model <- feols(
  scope1_emissions_log ~ treatment_exposed * post_announcement + log(total_assets) + roa + leverage + emissions_trajectory |
    `Identifier (RIC)` + year,
  data = df_panel %>% filter(scope1_reported == 1, scope1_emissions_log > 0),
  vcov = ~ `Identifier (RIC)`
)

summary(reported_model)
```

```{r modelled }
estimated_model <- feols(
  scope1_emissions_log ~ treatment_exposed * post_announcement + log(total_assets) + roa + leverage + emissions_trajectory |
    `Identifier (RIC)` + year,
  data = df_panel %>% filter(scope1_estimated == 1, scope1_emissions_log > 0),
  vcov = ~ `Identifier (RIC)`
)

summary(estimated_model)
```

## Industry Heterogeneity

```{r split sample models}
high_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    "+ log(total_assets) + roa + leverage + emissions_trajectory| ",
    "`Identifier (RIC)` + year"
  )),
  data = df_panel %>% filter(high_emission_sector == 1),
  vcov = ~ `Identifier (RIC)`
)

low_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    "+ log(total_assets) + roa + leverage + emissions_trajectory| ",
    "`Identifier (RIC)` + year"
  )),
  data = df_panel %>% filter(high_emission_sector == 0),
  vcov = ~ `Identifier (RIC)`
)

summary(high_model)
summary(low_model)
```

```{r triple difference model}
triple_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    "+ log(total_assets) + roa + leverage + emissions_trajectory| ",
    "`Identifier (RIC)` + year"
  )),
  data = df_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(triple_model)
```

```{r etable}
etable(
  high_model,
  low_model,
  triple_model,
  headers = c("Carbon-Intensive", "Other Sectors", "Pooled Triple-Diff"),
  drop = c("log(total_assets)", "roa", "leverage", "Constant"),
  keep = c("%treatment_exposedTRUE:post_announcementTRUE",
           "%treatment_exposedTRUE:post_announcementTRUE:high_emission_sector1"),
  se.below = TRUE,
  digits = 3
)
```

## Robustness Checks

```{r baseline model for robusntess}
summary(fe_model) #see cell named: "r base plus controlled plus fixed effects"
```

```{r uk firms as control group}
uk_control_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    "+ log(total_assets) + roa + leverage + emissions_trajectory| ",
    "`Identifier (RIC)` + year"
  )),
  data = df_panel %>%
    filter(treatment_exposed == TRUE | UK_based == TRUE), #filter the data panel to only include treatment exposed firms, and UK_based firms
  vcov = ~ `Identifier (RIC)`
)

summary(uk_control_model)
```

```{r placebo test with uk as treated group}
#uses the untreated UK group as the treated group to check for spillover/invalid info
df_placebo <- df_panel %>%
  filter(group %in% c("Placebo", "Control")) %>%
  mutate(fake_treatment = ifelse(group == "Placebo", 1, 0))

placebo_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    "+ log(total_assets) + roa + leverage + emissions_trajectory| ",
    "`Identifier (RIC)` + year"
  )),
  data = df_placebo,
  vcov = ~ `Identifier (RIC)`
)

summary(placebo_model)
```


```{r placebo treatment T-3}
#this model checks a different treatment year, in this case 2019 or time T - 3
#first we filter the dataset to exclude the treatment period
df_placebo <- df_panel %>%
  filter(year <= 2021)

#then we run the model
placebo_model_2019 <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    "+ log(total_assets) + roa + leverage + emissions_trajectory| ",
    "`Identifier (RIC)` + year"
  )),
  data = df_placebo,
  vcov = ~ `Identifier (RIC)`
)

summary(placebo_model_2019)
```

```{r no industry z-scoring}
#no clustering by the industry or year, just the raw score
no_cluster_model <- feols(
  as.formula(paste0(
    "greenwashing_esg_raw ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + emissions_trajectory | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(no_cluster_model)

#year-standardized greenwashing index
y_cluster_model <- feols(
  as.formula(paste0(
    "greenwashing_esg_y ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage  + emissions_trajectory| ",
    "`Identifier (RIC)` + year"
  )),
  data = df_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(y_cluster_model)
```

```{r winsorization }
#managing of extreme outliers, take the bottom and top 5% and set to the 5th percentile value, and 95th percentile respectively 

#function to winsorise
winsorize_5_95 <- function(x) {
  p5 <- quantile(x, 0.05, na.rm = TRUE)
  p95 <- quantile(x, 0.95, na.rm = TRUE)
  pmax(pmin(x, p95), p5)
}

df_panel <- df_panel %>%
  mutate(
    !!paste0(y_var, "_wins") := winsorize_5_95(.data[[y_var]])
  )

robustness_model_wins <- feols(
  as.formula(paste0(
    paste0(y_var, "_wins"), " ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + emissions_trajectory | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(robustness_model_wins)
```

# Difference-in-differences (with matching)

```{r propensity score matching}
df_baseline <- df_panel %>% filter(year == 2021) #filter to baseline year closest to intervention but pre-intervention

#generate the formula
psm_formula <- as.formula(paste("treatment ~ log(total_asset) + `GICS Sector Code` +", y_var)) #matching based on size, industry, and y_var

#run the matching
psm_match <- matchit(psm_formula, # matching based on size, industry, greenwashing index
                     data = df_baseline, 
                     method = "nearest", 
                     distance = "logit")

#get the list of matched Firm IDs
matched_ids <- match.data(psm_match)$firm_id
#filter the original panel data to only keep the matched firms
df_psm_panel <- df_panel %>% filter(firm_id %in% matched_ids)
```

```{r Coarsened exact matching}
#build the formula dynamically
cem_formula <- as.formula(paste("treatment ~ log(total_asset) + `GICS Sector Code` +", y_var))
#match based on the formula
cem_match <- matchit(cem_formula, 
                     data = df_baseline, 
                     method = "cem", 
                     k2k = TRUE) # 'k2k = TRUE' ensures 1-to-1 matching (one control for one treatment)
#get matched IDS
matched_ids_cem <- match.data(cem_match)$firm_id
#filter the data
df_cem_panel <- df_panel %>% filter(firm_id %in% matched_ids_cem)
```

