---
title: "Analysis"
author: "Melle Stephen Oudshoorn"
date: "`r Sys.Date()`"
output: html_document
---

# Setup and work

```{r setup chunk, echo = TRUE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
) #setup standards for Rmarkdown

set.seed(19) #set random seed to make sure random results are reproducable if applicable
```

```{r load packages and key lists}
#basic stuff
library(tidyverse) #basic functions for filtering and mutation
#stuff for economic modeling
library(modelsummary) #to get summaries for complicated models
library(fixest) #runs fast FE regressions
library(MatchIt) #for matching algorithms to balance groups
library(car) #for diagnostic plots and testing
library(psych) # descriptive statistics
#pretty tables
library(knitr) #formatted Markdown tables with kable
library(gt) #builds presentation ready tables
library(kableExtra) #adds formatting options to kable
#data visualisation
library(patchwork) #simplifies seperate plots into single multi-panel layouts
library(scales) #handles axis scaling, number formatting, and coloring
library(extrafont) #fonts (for Times New Roman)
library(ggplot2) #PLOTTING YIPPIEEE

eu_countries <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czech Republic",
  "Denmark", "Estonia", "Finland", "France", "Germany", "Greece", "Hungary",
  "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Malta", "Netherlands",
  "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", "Spain", "Sweden"
) #I forgot to import variable whether companies are in EU so I can use these

y_var <- "greenwashing_esg" #dependent variable, done like this so I can check alternative versions of greenwashing very quickly.

df_panel <- read_csv("panel_data_lseg.csv") #import the data panel that was gathered from Refintiv Eikon (LSEG)
```


```{r describe the disclosure variables}
all_disclosure_vars <- c("esg_disclosure_score", "environmental_disclosure_score", "social_disclosure_score", "governance_disclosure_score")
#describe the disclosure scores for these independent pillars
psych::describe(df_panel %>%
  select(all_of(all_disclosure_vars))) #describe them in a table

df_esg_disclosure_long <- df_panel %>%
  select(all_of(all_disclosure_vars)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  ) #prep data for a plot

ggplot(df_esg_disclosure_long, aes(x = value)) + #run a plot for each variable
  geom_histogram(bins = 50) +
  facet_wrap(~ variable, scales = "free") + #facet wrap to put them all in one plot
  labs(
    title = "Distribution of ESG Variables",
    x = NULL,
    y = "Count"
  ) +
  theme_minimal()
```

```{r boxplots of disclosure and performance score to compare their base scores}
df_panel %>%
  pivot_longer(c(esg_score, esg_disclosure_score)) %>% #pivot the table
  ggplot(aes(x = name, y = value)) +
  geom_boxplot(fill = "#b4bfff", color = "#5874ff", width = 0.4) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  theme_classic() +
  theme(axis.title = element_blank())

#shows that standardising is necessary
```

```{r computing the greenwashing index according to yu et al.}
df_panel <- df_panel %>%
  group_by(year, `GICS Sector Code`) %>%  #group by year and sector to remove mechanical year and industry variation
  mutate(
    #standardise the disclosure scores within industry-year groups
    esg_disclosure_z = as.numeric(scale(esg_disclosure_score)),
    environmental_disclosure_z = as.numeric(scale(environmental_disclosure_score)),
    social_disclosure_z = as.numeric(scale(social_disclosure_score)),
    governance_disclosure_z = as.numeric(scale(governance_disclosure_score)),
    #standardise performance scores within their industry-year groups
    esg_performance_z = as.numeric(scale(esg_score)),
    environmental_performance_z = as.numeric(scale(environmental_pillar_score)),
    social_performance_z = as.numeric(scale(social_pillar_score)),
    governance_performance_z = as.numeric(scale(governance_pillar_score)),
    
    #compute greenwashing indeces (standardised disclosure - standardised performance)
    
    #baseline greenwashing index
    greenwashing_esg = esg_disclosure_z - esg_performance_z,
    
    #pillared greenwashing indeces
    greenwashing_environmental = environmental_disclosure_z - environmental_performance_z,
    greenwashing_social = social_disclosure_z - social_performance_z,
    greenwashing_governance = governance_disclosure_z - governance_performance_z,
    
    #robustness check
    greenwashing_environmental_alt = esg_disclosure_z - environmental_performance_z #robustness check (unused as not really a change)
  ) %>% 
  ungroup()

#full sample alternative (used forrobustness)
df_panel <- df_panel %>%
  mutate(
    esg_disclosure_z_raw = as.numeric(scale(esg_disclosure_score)),
    esg_performance_z_raw = as.numeric(scale(esg_score)),
    greenwashing_esg_raw = esg_disclosure_z_raw - esg_performance_z_raw
  )

#year clustered alternative (used forrobustness)
df_panel <- df_panel %>%
  group_by(year) %>%
  mutate(
    esg_disclosure_z_y = as.numeric(scale(esg_disclosure_score)),
    esg_performance_z_y = as.numeric(scale(esg_score)),
    greenwashing_esg_y = esg_disclosure_z_y - esg_performance_z_y
  ) %>%
  ungroup()

#industry clustered alternative (used for robustness)
df_panel <- df_panel %>% 
  group_by(`GICS Sector Code`) %>%
  mutate(
    esg_disclosure_z_i = as.numeric(scale(esg_disclosure_score)),
    esg_performance_z_i = as.numeric(scale(esg_score)),
    greenwashing_esg_i = esg_disclosure_z_i - esg_performance_z_i
  ) %>%
  ungroup()
```

## Computing the matched DF's 

```{r propensity score matching}
covariates <- c("log_total_assets","roa","leverage","tobins_q","share_turnover","gics_sector_code")

df_baseline <- df_panel %>%
  filter(year == 2019) %>% 
  mutate(
    log_total_assets = ifelse(
      !is.na(total_assets) & total_assets > 0,
      log(total_assets),
      NA_real_
    ),
    gics_sector_code = factor(`GICS Sector Code`)
  ) %>%
  #only keep firms that have all must‑have covariates non‑missingtomake sure the panel is matchable
  filter(
    across(all_of(covariates), ~ !is.na(.x)),
    !is.na(treatment_exposed),
    !is.na(`Country of Incorporation`),
    !is.na(gics_sector_code)
  ) %>%
  distinct(`Identifier (RIC)`, .keep_all = TRUE)

#define the formula
psm_formula <- as.formula(
  paste("treatment_exposed ~",
        paste(covariates, collapse = " + "))
)

#do the genuine matching
psm_match <- matchit(
  formula = psm_formula, #insert formula
  data = df_baseline, #use the dataset
  method = "nearest",
  distance = "logit"
)

#get the matched firm IDs
matched_ids <- match.data(psm_match) %>%
  pull(`Identifier (RIC)`) %>%
  unique()

#create the panel with the matched IDs
df_psm_panel <- df_panel %>%
  filter(`Identifier (RIC)` %in% matched_ids)

summary(psm_match)
```

```{r mahalanobis matching}
#perform the mahalanbis matching (MHM)
mdm_match <- matchit(psm_formula,
                     data = df_baseline,
                     method = "nearest",
                     distance = "mahalanobis")

matched_ids_mdm <- match.data(mdm_match)$`Identifier (RIC)` #get the matched IDs
df_mdm_panel <- df_panel %>% filter(`Identifier (RIC)` %in% matched_ids_mdm) #create the panel based on the matched IDs

summary(mdm_match) #summarise
```


```{r coarsened exact matching}
#match based on the formula
cem_match <- matchit(psm_formula, 
                     data = df_baseline, 
                     method = "cem",
                     cutpoints = list(
                       "log(total_assets)" = 15, # Break assets into 15 tight groups
                       "roa" = 10,               # Break ROA into 10 tight groups
                       "leverage" = 10           # Break leverage into 10 tight groups
                     ))
#get matched IDS
matched_ids_cem <- match.data(cem_match)$`Identifier (RIC)`
#filter the data to create the matched panel
df_cem_panel <- df_panel %>% 
  filter(`Identifier (RIC)` %in% match.data(cem_match)$`Identifier (RIC)`)

summary(cem_match)
```

```{r choosing the dynamic dataset for analysis}
df_analysis <- df_panel #redundant now, but used to be done otherwise.
```

# Standard descriptive statistics

## Sample structure

```{r basic panel structure}
# number of firms, years, observations in the full sample
df_analysis %>% 
  summarise(N_obs = comma(n()), N_firms = comma(n_distinct(`ISIN Code`)), Years = n_distinct(year), N_countries = comma(n_distinct(`Country of Incorporation`))) %>%
  kable(booktabs = TRUE) %>%
  kable_styling() #make it pretty


# observations per year for the full sample
df_analysis %>% count(year) %>% mutate(Share = percent(n/sum(n))) %>%select(Year=year, Obs=n, Share) %>% kable(booktabs=T) %>% kable_styling()
```

```{r N observations by country treatment status and GICS sector}
#geographic table, where you count the observations per country
geo_table <- df_analysis %>% #initialise
  group_by(Country = `Country of Incorporation`) %>% #group by the country of incorporation
  summarise(Obs = n(), `Unique Firms` = n_distinct(`ISIN Code`), .groups = "drop") %>% #summarise the number of unique firms per country
  arrange(desc(Obs)) %>% #high to low
  mutate(Share = percent(Obs / sum(Obs))) %>% #calc the perc of the sample
  select(Country, Obs, `Unique Firms`, Share) %>% #get the coefficients of interest
  #format
  kable(format = "html", booktabs = TRUE, caption = "Geographic Sample Distribution and Composition") %>% 
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE, font_size = 14) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#444444")

save_kable(geo_table, file = "geographic_distribution.png", zoom = 2) #save itt
geo_table #display it

#show the share per group
df_analysis %>% 
  count(Group = group, name = "Obs") %>%  #count by group (treatment/control/placebo)
  mutate(Share = scales::percent(Obs / sum(Obs))) %>% #calc percentage 
  #format
  kable(booktabs = TRUE) %>% 
  kable_styling(latex_options = "striped")

#show the distribution of industry
df_analysis %>% 
  filter(`GICS Sector Name` != "Financials") %>% #filter out the GICS sector name
  count(Sector = `GICS Sector Name`, sort = TRUE, name = "Obs") %>%  #count the number of variables per sector
  mutate(Share = percent(Obs / sum(Obs))) %>% #calc the share
  #format
  kable(booktabs = TRUE) %>% 
  kable_styling()

#distribution of group and year while also showing perc's
df_analysis %>% 
  count(Year = year, treatment_exposed) %>%  #count observations per year per group
  pivot_wider(names_from = treatment_exposed, values_from = n) %>% #pivor]t
  mutate(Pct_Treated = scales::percent(`TRUE` / (`FALSE` + `TRUE`))) %>%
  select(Year, Control = `FALSE`, Treated = `TRUE`, Pct_Treated) %>%
  #format
  kable(digits = 0, booktabs = TRUE) %>% 
  kable_styling()
```

```{r by treatment control and placebo firms by region}
#create a subsample for each region
df_region <- df_panel %>%
  distinct(`Identifier (RIC)`, .keep_all = TRUE) %>% #get all distinct companies
  mutate( #create the region variable
    region = case_when(
      `Country of Incorporation` %in% c(
        "Austria","Belgium","Cyprus","Denmark","Finland","France","Germany",
        "Greece","Hungary","Ireland","Italy","Luxembourg","Malta",
        "Netherlands","Poland","Portugal","Romania","Slovenia","Spain","Sweden"
      ) ~ "EU",
      
      `Country of Incorporation` %in% c(
        "United Kingdom","Norway","Switzerland","Iceland"
      ) ~ "UK & Europe non-EU",
      
      `Country of Incorporation` %in% c(
        "United States","Canada"
      ) ~ "North America",
      
      `Country of Incorporation` %in% c(
        "China (Mainland)","Hong Kong","India","Indonesia","Japan","Malaysia",
        "Pakistan","Philippines","Singapore","South Korea","Taiwan","Thailand",
        "Vietnam","Sri Lanka","Kazakhstan","New Zealand"
      ) ~ "Asia-Pacific",
      
      `Country of Incorporation` %in% c(
        "Israel","Kuwait","Qatar","Saudi Arabia","United Arab Emirates","Bahrain","Oman","Turkey"
      ) ~ "Middle East",
      
      `Country of Incorporation` %in% c(
        "Nigeria","South Africa","Kenya","Egypt","Morocco"
      ) ~ "Africa",
      
      `Country of Incorporation` %in% c(
        "Argentina","Brazil","Chile","Colombia","Mexico","Peru","Panama"
      ) ~ "Latin America",
      
      `Country of Incorporation` %in% c(
        "Cayman Islands","Jersey","Bermuda","Gibraltar","Guernsey","Faroe Islands",
        "Isle of Man","Marshall Islands","Liberia","British Virgin Islands","Curacao",
        "Mauritius"
      ) ~ "Offshore / other"
    )
  )

#format the distribution in a ggplot
p_region_distribution <- df_region %>%
  filter(!is.na(`Country of Incorporation`)) %>% 
  count(region, group) %>% #count the number of firms for each group per region
  mutate(
    region = fct_reorder(region, n, .fun = sum),
    group = factor(group, levels = c("Control", "Placebo", "Treatment"))
  ) %>%
  ggplot(aes(x = region, y = n, fill = group)) +
  geom_col(width = 0.65) + 
  coord_flip() +
  #formatting
  scale_fill_manual(values = c("Control"   = "#C0392B","Placebo"   = "#7F8C8D", "Treatment" = "#108A99"   )) +
  labs(x = NULL,y = "Number of firms",fill = NULL) +
  theme_minimal(base_size = 12, base_family = "Times New Roman") +
  theme(
    axis.title.x       = element_text(margin = margin(t = 10), size = 11, face = "plain"),
    axis.text.y        = element_text(size = 11, color = "black"),
    axis.text.x        = element_text(size = 10, color = "black"),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(color = "#E5E7E9", linewidth = 0.4), 
    panel.grid.minor.x = element_blank(),
    legend.position    = "top",
    legend.justification = "left",
    legend.text        = element_text(size = 11),
    legend.margin      = margin(b = -5, l = -10)
  )

p_region_distribution #display the plot

#save the plot to output folder
if(!dir.exists("output")) dir.create("output") #create the folder if it doesnt exist yet
ggsave(
  plot     = p_region_distribution,
  filename = "output/region_distribution.png", 
  type     = "cairo", 
  width    = 7, 
  height   = 4.5,
  dpi      = 300 
)
```

```{r kernal density plot of GW by treatment PRE/POST}
#create a new df where you have a pre and post variable
df_density <- df_analysis %>%
  mutate(
    period = if_else(year <= 2021, "Pre-treatment (2017–2021)", "Post-treatment (2022–2025)"),
    group = if_else(treatment_exposed == 1, "Treated Group", "Control Group")
  )

ggplot(df_density, aes(x = .data[[y_var]], fill = period, color = period)) +
  geom_density(alpha = 0.25, linewidth = 1, adjust = 1) +
  facet_wrap(~ group) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "grey50") +
  labs(
    title = "Kernel density of Greenwashing Index before and after CSRD",
    x = "Greenwashing Index",
    y = "Density",
    fill = NULL,
    color = NULL
  ) +
  scale_fill_manual(values = c("Pre-treatment (2017–2021)" = "#1f77b4",
                               "Post-treatment (2022–2025)" = "#C0392B")) +
  scale_color_manual(values = c("Pre-treatment (2017–2021)" = "#1f77b4",
                                "Post-treatment (2022–2025)" = "#C0392B")) +
  geom_vline(xintercept = 0, linetype = "dotted", color = "black") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Descriptive statistics and correlations

```{r descriptive statistics of points of interest}
#define variables and formal names for the table
descr_vars <- c(
  "total_assets"               = "Total Assets (Billion EUR)",
  "roa" = "Return on Assets (net income / total assets)",
  "leverage" = "Financial Leverage (Total debt / Total assets)",
  "share_turnover" = "share turnover / market capitalisation",
  "tobins_q" = "Market Valuation (Market capitalisation / total assets)"
)
descr_vars <- c(setNames("Greenwashing Index", y_var), descr_vars)

#build the table
desc_table <- map_dfr(names(descr_vars), function(v) {
  x <- df_analysis[[v]]
  
  #scale assets to Billions only for display
  if(v == "total_assets") x <- x / 1e9
  data.frame(
    Variable = descr_vars[v],
    N        = sum(!is.na(x)),
    Mean     = mean(x, na.rm = TRUE),
    St.Dev.  = sd(x, na.rm = TRUE),
    Min      = min(x, na.rm = TRUE),
    P25      = quantile(x, 0.25, na.rm = TRUE),
    Median   = median(x, na.rm = TRUE),
    P75      = quantile(x, 0.75, na.rm = TRUE),
    Max      = max(x, na.rm = TRUE)
  )
})

#render the table
desc_table %>%
  kable(format = "html", digits = 3, caption = "Table X: Descriptive Statistics", 
        format.args = list(big.mark = ","),
        col.names = c("Variable", "N", "Mean", "St. Dev.", "Min", "P25", "Median", "P75", "Max")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#444444")
```

## Pre-treatment Parallel trends

```{r distribution of greenwashing pre announcement}
#new df with pre-treatment only
df_descriptive <- df_panel %>%
  filter(year < 2022) %>%     # pre‑treatment
  select(`Identifier (RIC)`, year, group, .data[[y_var]]) %>% #select relevant columns
  drop_na(.data[[y_var]]) #drop the NAs for GW score

#create a plot to compare the distributions and means for the different groups
p_dist <- df_descriptive %>%
  mutate(group = fct_relevel(group, "Control", "Placebo", "Treatment")) %>%
  ggplot(aes(x = group, y = .data[[y_var]], fill = group)) +
  geom_violin(trim = FALSE, alpha = 0.8, color = "black", linewidth = 0.3) + #violin plot
  geom_boxplot(width = 0.18, outlier.alpha = 0.3,fill = "white", color = "black",linewidth = 0.6) +   #boxplot overlay timeee
  #formatting
  scale_fill_manual(values = c("Control"= "#C0392B","Placebo"= "#7F8C8D","Treatment" = "#108A99"),guide = "none") +
  labs(x = NULL,y = "Greenwashing Index (ESG-based)" ) +
  theme_minimal(base_size = 12, base_family = "Times New Roman") +
  theme(
    axis.title.y       = element_text(margin = margin(r = 10), size = 11),
    axis.text.x        = element_text(size = 11, color = "black"),
    axis.text.y        = element_text(size = 10, color = "black"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "#E5E7E9", linewidth = 0.4),
    panel.grid.minor.y = element_blank()
  )

p_dist  #display plot

ggsave( #save plot to output folder
  plot     = p_dist,
  filename = "output/greenwashing_distribution.png", 
  type     = "cairo",  
  width    = 7, 
  height   = 4.5,
  dpi      = 300       
)
```

```{r balance test}
# Turn off scientific notation globally for this chunk
options(scipen = 999)

# filter for pre-treatment years
df_pre <- df_analysis %>%
  filter(year >= 2019 & year <= 2021) %>%
  filter(group %in% c("Treatment", "Control"))

table_3_data <- map_dfr(names(descr_vars), function(v) {
  
  # Conduct t-test: variable ~ group
  t_res <- t.test(df_pre[[v]] ~ df_pre$group)
  
  # Extract means (Note: R orders groups alphabetically)
  means <- t_res$estimate
  p_val <- t_res$p.value
  
  # Format p-value cleanly for an academic table
  p_val_formatted <- if(p_val < 0.001) "< 0.001" else as.character(round(p_val, 3))
  
  data.frame(
    Variable = descr_vars[v],
    `Treatment Mean` = round(means[2], 3), 
    `Control Mean` = round(means[1], 3),
    `Diff (T-C)` = round(means[2] - means[1], 3),
    `p-value` = p_val_formatted,
    check.names = FALSE # Prevents R from converting spaces to dots in column names
  )
})

table_3_data %>%
  kable(
    format = "html",
    align = c("l", "r", "r", "r", "r"),
    caption = "Table 3: Pre-Treatment Balance Test (2019–2021 Means)",
    col.names = c("Variable", "Treatment Mean", "Control Mean", "Diff. (T-C)", "p-value")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#444444") %>%
  footnote(general = "p-values are obtained from two-sample Welch t-tests comparing the Treatment and Control groups during the 2019–2021 baseline window.")
```

```{r parallel trends plot with controls}
#filter the data for the event study
df_event <- df_analysis %>% 
  filter(group %in% c("Treatment", "Control")) %>%
  filter(!is.na(roa), !is.na(leverage), !is.na(total_assets), !is.na(`GICS Sector Code`), !is.na(share_turnover), !is.na(tobins_q)) #filter across NA's for the control variables

event_study <- feols( #run the event study and calc point estimates
  as.formula(paste0(
    y_var, " ~ i(year, treatment_exposed, ref = 2021)", 
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q | `Identifier (RIC)` + year"
  )),
  data = df_event,
  vcov = ~ `Identifier (RIC)`
)

#save the plot to the output folder
png("output/event_study_results.png", width = 7.5, height = 4.8, units = "in", res = 300)
iplot(
  event_study,
  main = "", 
  xlab = "Year",
  ylab = "Estimated Effect (Coefficient)"
)
dev.off() #close the file saver

#print the plot in RSTUDIO output
iplot(
  event_study,
  main = "Event Study: Impact of CSRD on Greenwashing", 
  xlab = "Year",
  ylab = "Estimated Effect (Coefficient)"
)
```


```{r parallel trends plot no controls}
#do the same thing but with no controls
df_event <- df_analysis %>% 
  filter(group %in% c("Treatment", "Control"))

# run the unconditional model
event_study_perf_uncond <- feols(
  as.formula(paste0(
     y_var, " ~ i(year, treatment_exposed, ref = 2021)", 
    " | `Identifier (RIC)` + year"
  )),
  data = df_event,
  vcov = ~ `Identifier (RIC)`
)

#save the clean plot to the output folder
png("output/event_study_performance_uncond.png", width = 7.5, height = 4.8, units = "in", res = 300)
iplot(
  event_study_perf_uncond,
  main = "",
  xlab = "Year",
  ylab = "Estimated Effect (Coefficient)"
)
dev.off()

#view it in Rstudio
iplot(
  event_study_perf_uncond,
  main = "Unconditional Event Study: Impact of CSRD on standardised ESG performance score", 
  xlab = "Year",
  ylab = "Estimated Effect (Coefficient)"
)
```

### Event study decoupled GW index

```{r performance z parallel trends plot with controls}
#do the same but with disclosure score as the plot
df_event <- df_analysis %>% 
  filter(group %in% c("Treatment", "Control")) %>%
  filter(!is.na(roa), !is.na(leverage), !is.na(total_assets), !is.na(`GICS Sector Code`))

event_study_disc_cond <- feols(
  as.formula(paste0(
    "esg_disclosure_z ~ i(year, treatment_exposed, ref = 2021)", 
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q | `Identifier (RIC)` + year"
  )),
  data = df_event,
  vcov = ~ `Identifier (RIC)`
)

#save conditional version in output folder regarding the disclosure score
png("output/event_study_disclosure_cond.png", width = 7.5, height = 4.8, units = "in", res = 300)
iplot(event_study_disc_cond,main = "",xlab = "Year",ylab = "Estimated Effect (Coefficient)")
dev.off()

# View it instantly on your screen right now
iplot(event_study_disc_cond,main = "Conditional Event Study: Impact of CSRD on standardised ESG disclosure score", xlab = "Year",ylab = "Estimated Effect (Coefficient)")
```

```{r disclosure z parallel trends plot with controls}
#once more but for esg performance z score
event_study_perf_cond <- feols(
  as.formula(paste0(
    "esg_performance_z ~ i(year, treatment_exposed, ref = 2021)", 
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q | `Identifier (RIC)` + year"
  )),
  data = df_event,
  vcov = ~ `Identifier (RIC)`
)

#save to output folder
png("output/event_study_performance_cond.png", width = 7.5, height = 4.8, units = "in", res = 300)
iplot(
  event_study_perf_cond,
  main = "", # Title handled by LaTeX caption
  xlab = "Year",
  ylab = "Estimated Effect (Coefficient)"
)
dev.off()

#view
iplot(
  event_study_perf_cond,
  main = "Conditional Event Study: Impact of CSRD on standardised ESG performance score", 
  xlab = "Year",
  ylab = "Estimated Effect (Coefficient)"
)
```


```{r parallel trends visualisation}
p_trend <- df_analysis %>%
  filter(year >= 2017) %>% #filter is kinda irrelevant right now 
  mutate(Group = factor(treatment_exposed, 
                        levels = c("FALSE", "TRUE"), labels = c("Control Group", "Treatment Group"))) %>% #convert binary into labels
  group_by(year, Group) %>% #group it
  summarise(mean_gw = median(.data[[y_var]], na.rm = TRUE), .groups = "drop") %>% #do the mean per year
  ggplot(aes(x = year, y = mean_gw, color = Group, group = Group)) + #plot the greenwashing over time
  #add the strructural rules
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  #add event miilesntones
  geom_vline(xintercept = 2022, linetype = "dashed", color = "#5D6D7E", linewidth = 0.5) +
  geom_vline(xintercept = 2024, linetype = "dashed", color = "#5D6D7E", linewidth = 0.5) +
  #text annotations for milestones
  annotate("text", x = 2021.8, y = 0.12, label = "CSRD Announcement", 
           angle = 90, family = "Times New Roman", size = 3.5, color = "#2C3E50", hjust = 1) +
  annotate("text", x = 2023.8, y = 0.12, label = "CSRD Effective Date", 
           angle = 90, family = "Times New Roman", size = 3.5, color = "#2C3E50", hjust = 1) +
  
  #formatting
  scale_color_manual(values = c("Control Group" = "#C0392B", "Treatment Group" = "#108A99")) +
  scale_x_continuous(breaks = seq(2017, 2025, by = 2)) +
  labs(
    x = "Fiscal Year",
    y = "Average Greenwashing Index"
  ) +
  #academic layout
  theme_minimal(base_size = 12) +
  theme(
    legend.title       = element_blank(),
    legend.position    = "bottom",
    axis.title.x       = element_text(margin = margin(t = 10)),
    axis.title.y       = element_text(margin = margin(r = 10)),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "#E5E7E9", linewidth = 0.4),
    panel.grid.minor.y = element_blank()
  )

p_trend #display plot

#save the plot to output folder
ggsave(
  plot     = p_trend,
  filename = "output/greenwashing_trends.png",
  width    = 7.5,
  height   = 4.8,
  dpi      = 300
)
```

```{r decomposed time trend}
df_trends_long <- df_panel %>%
  filter(year >= 2017) %>% #filter kinda irrelevant
  group_by(year, treatment_exposed) %>%
  summarise(
    "ESG Disclosure Score" = mean(esg_disclosure_z, na.rm = TRUE), #take the mean values of each year
    "ESG Performance Score" = mean(esg_performance_z, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  #pivot only the two components into long format
  pivot_longer(
    cols = c("ESG Disclosure Score", "ESG Performance Score"),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  mutate(
    group_label = ifelse(treatment_exposed == TRUE, "Treatment", "Control")
  )
#plot these two over time
ggplot(
  df_trends_long,
  aes(x = year, y = Value, color = group_label, linetype = Metric, group = interaction(group_label, Metric))
) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  
  #policy lines for CSRD announcement and effect
  geom_vline(xintercept = c(2022, 2024), linetype = "dashed", color = "gray50") +
  #labels for the events
  annotate("text", x = 2022, y = Inf, label = "CSRD announcement", angle = 90, vjust = 1.2, hjust = 1.1, size = 3.5, color = "gray30") +
  annotate("text", x = 2024, y = Inf, label = "CSRD goes in effect", angle = 90, vjust = 1.2, hjust = 1.1, size = 3.5, color = "gray30") +
  #color scheme
  scale_color_manual(values = c("Control"   = "#C0392B","Treatment" = "#108A99")) +
  scale_linetype_manual(values = c("ESG Disclosure Score"  = "solid", "ESG Performance Score" = "dashed")) +
  #add the titles and labels
  labs(
    title = "ESG Disclosure and Performance Trends over Time",
    subtitle = "Median standardised scores by treatment status",
    x = "Year",
    y = "Median Standardised Score",
    color = NULL,
    linetype = NULL
  ) +
  # theming
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
    panel.grid.minor = element_blank()
  )
```

# Difference-in-differences (regular type)

## H1 Main effect

```{r base model}
base_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement" #base, only the announcement and treatment exposure
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(base_model)
```

```{r base plus controls}
controls_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q " #include the controls
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(controls_model)
```

```{r base plus controlled plus firm fixed effects}
firm_fe_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q  |", #controls
    "`Identifier (RIC)`" #fixed effects for firm
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(firm_fe_model)
```

```{r base plus controls plus time fixed effects}
time_fe_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q  |", #controls
    "year" #fixed effects for year
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(time_fe_model)
```

```{r base plus controlled plus full fixed effects}
fe_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q  |", #controls
    "`Identifier (RIC)` + year" #fixed effects for firm and year
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model)
```

```{r main effect post enforcement}
fe_model_enforcement <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_csrd", #base
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q  |", #controls
    "`Identifier (RIC)` + year" #fixed effects for firm and year
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_enforcement)
```


## H2 Mechanism test

```{r the narrative mechanism}
#dependent variable is the standardised disclosure score (use most complete specification)
mechanism_disclosure <- feols(
  esg_disclosure_z ~ treatment_exposed * post_announcement + log(total_assets) + roa + leverage + share_turnover + tobins_q |
    `Identifier (RIC)` + year,
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(mechanism_disclosure)
```

```{r the performance mechanism}
#dependent variable is the standardised performance score (use most complete specification)
mechanism_performance <- feols(
  esg_performance_z ~ treatment_exposed * post_announcement + log(total_assets) + roa + leverage + share_turnover + tobins_q  |
    `Identifier (RIC)` + year,
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(mechanism_performance)
```

## H3 Industry Heterogeneity

```{r triple d formula}
#define the formula
triple_d_formula <- as.formula(paste0(
  y_var, " ~ treatment_exposed * post_announcement * low_emission_sector + ",
  "log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
  "`Identifier (RIC)` + year"
))

#run the fixed-effects OLS
triple_d <- feols(triple_d_formula,data = df_analysis,vcov = ~ `Identifier (RIC)`)

summary(triple_d)
```

```{r seperated heterogeneity}
#create two subsamples:
df_low_intensity_sample <- df_analysis %>% filter(low_emission_sector == TRUE)
df_high_intensity_sample <- df_analysis %>% filter(low_emission_sector == FALSE)

#run the main regression for each subsample
low_intensity_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q  |", #controls
    "`Identifier (RIC)` + year" #fixed effects for firm and year
  )),
  data = df_low_intensity_sample,
  vcov = ~ `Identifier (RIC)`
)

high_intensity_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement", #base
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q  |", #controls
    "`Identifier (RIC)` + year" #fixed effects for firm and year
  )),
  data = df_high_intensity_sample,
  vcov = ~ `Identifier (RIC)`
)

#view the results
summary(low_intensity_model)
summary(high_intensity_model)
```




```{r sector heterogeneity plot}
#extraction pipeline
sectors <- unique(df_analysis$`GICS Sector Name`) #get all sectors
sectors <- sectors[!is.na(sectors)]

#run the model for every sector
df_gics_quick <- map_dfr(sectors, function(sec) { 
  df_sub <- df_analysis %>% filter(`GICS Sector Name` == sec)
  
  tryCatch({
    model <- feols(
      as.formula(paste0(
        y_var, " ~ treatment_exposed * post_announcement + ",
        "log(total_assets) + roa + leverage + share_turnover + tobins_q | `Identifier (RIC)` + year"
      )),
      data = df_sub,
      vcov = ~ `Identifier (RIC)`
    )
    
    ct <- as.data.frame(coeftable(model))
    ci <- as.data.frame(confint(model, level = 0.95))
    row_name <- "treatment_exposedTRUE:post_announcementTRUE"
    
    tibble(
      sector = sec,
      estimate = ct[row_name, 1],
      conf.low = ci[row_name, 1],
      conf.high = ci[row_name, 2]
    )
  }, error = function(e) {
    return(NULL) 
  })
})

#relabel 
high_emission_macro <- c("Energy", "Materials", "Utilities")

#plot these sectors for the high emissions
df_gics_plot <- df_gics_quick %>%
  mutate(
    is_high_emission = ifelse(sector %in% high_emission_macro, "High-Emission Sectors", "Other Sectors")
  )

#create ggplot 
p_sector <- ggplot(df_gics_plot, aes(x = estimate, y = reorder(sector, estimate), color = is_high_emission)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.8) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), size = 0.6, linewidth = 0.9) +
  #color scheme
  scale_color_manual(values = c("High-Emission Sectors" = "#C0392B", "Other Sectors" = "#7F8C8D")) +
  #formatting
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.margin = margin(15, 15, 15, 15)
  ) +
  labs(
    x = "Estimated Treatment Effect (ATT)",
    y = NULL
  )

#save to output folder
ggsave("output/csrd_effect_by_sector.png", plot = p_sector, 
       width = 7.5, height = 4.8, units = "in", dpi = 300)

#print
print(p_sector)
```


## Robustness Checks

### Baseline robustness

```{r baseline model for robusntess}
summary(fe_model) #see cell named: "r base plus controlled plus fixed effects"
```

### Placebo tests

```{r uk firms as control group}
uk_control_model <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    "+ log(total_assets) + roa + leverage + share_turnover + tobins_q | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis %>%
    filter(treatment_exposed == TRUE | UK_based == TRUE), #filter the data panel to only include treatment exposed firms, and UK_based firms
  vcov = ~ `Identifier (RIC)`
)

summary(uk_control_model)
```

```{r placebo test with uk as treated group}
#uses the untreated UK group as the treated group to check for spillover/invalid info
df_placebo_uk <- df_analysis %>%
  filter(group %in% c("Placebo", "Treatment")) %>%
  mutate(
    # Ensure this is a factor or numeric for the regression
    fake_treatment = ifelse(group == "Placebo", 1, 0),
    post_val = ifelse(post_announcement == TRUE, 1, 0)
  )

placebo_model <- feols(
  as.formula(paste0(y_var, " ~ fake_treatment * post_val + log(total_assets) + leverage + share_turnover + tobins_q  | `Identifier (RIC)` + year")),
  data = df_placebo_uk,
  vcov = ~ `Identifier (RIC)`
)

summary(placebo_model)
```


```{r placebo treatment T-3}
#this model checks a different treatment year, in this case 2019 or time T - 3
#first we filter the dataset to exclude the treatment period
df_placebo_t3 <- df_analysis %>%
  filter(year < 2022) %>% 
  mutate(fake_post_2019 = ifelse(year >= 2019, 1, 0)) #create fake post period

#then we run the model
placebo_model_2019 <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * fake_post_2019", 
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_placebo_t3,
  vcov = ~ `Identifier (RIC)`
)

summary(placebo_model_2019)
```

### Alternative specifications for the Greenwashing index

```{r no industry z-scoring}
#no clustering by the industry or year, just the raw score
no_cluster_model <- feols(
  as.formula(paste0(
    "greenwashing_esg_raw ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(no_cluster_model)

#year-standardised greenwashing index
y_cluster_model <- feols(
  as.formula(paste0(
    "greenwashing_esg_y ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + share_turnover + tobins_q | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(y_cluster_model)

#industry_standardised greenwashing index
i_cluster_model <- feols(
  as.formula(paste0(
    "greenwashing_esg_i ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + share_turnover + tobins_q | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(i_cluster_model)
```

### Sensitivity to outlier handling

```{r winsorisation }
#managing of extreme outliers, take the bottom and top 5% and set to the 5th percentile value, and 95th percentile respectively 

#function to winsorise
winsorise_5_95 <- function(x) {
  p5 <- quantile(x, 0.05, na.rm = TRUE)
  p95 <- quantile(x, 0.95, na.rm = TRUE)
  pmax(pmin(x, p95), p5)
}

df_analysis <- df_analysis %>%
  mutate(
    !!paste0(y_var, "_wins") := winsorise_5_95(.data[[y_var]]) #winsorised for y_var
  )

robustness_model_wins <- feols(
  as.formula(paste0(
    paste0(y_var, "_wins"), " ~ treatment_exposed * post_announcement +",
    "log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_analysis,
  vcov = ~ `Identifier (RIC)`
)

summary(robustness_model_wins)
```

### non-parametric matching models

```{r psm model with fixed effects and controls}
fe_model_psm <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_psm_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_psm)
```

```{r psm model with fixed effects}
fe_model_psm_no_controls <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    " | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_psm_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_psm_no_controls)
```

```{r mdm model with fixed effects and controls}
fe_model_mdm <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_mdm_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_mdm)
```

```{r mdm model with fixed effects}
fe_model_mdm_no_controls <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    " | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_mdm_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_mdm_no_controls)
```

```{r cem model with fixed effects and controls}
fe_model_cem <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    " + log(total_assets) + roa + leverage + share_turnover + tobins_q  | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_cem_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_cem)
```

```{r}
fe_model_cem_no_controls <- feols(
  as.formula(paste0(
    y_var, " ~ treatment_exposed * post_announcement",
    " | ",
    "`Identifier (RIC)` + year"
  )),
  data = df_cem_panel,
  vcov = ~ `Identifier (RIC)`
)

summary(fe_model_cem_no_controls)
```

```{r summarising table}
robustness_list <- list(
  "Sophisticated model" = fe_model,
  "UK Control" = uk_control_model,
  "Placebo (UK)" = placebo_model,
  "Placebo (T-3)" = placebo_model_2019,
  "Raw GW" = no_cluster_model,
  "Year clustered GW" = y_cluster_model,
  "Industry clustered GW" = i_cluster_model,
  "Winsorised" = robustness_model_wins,
  "Propensity score matching" = fe_model_psm,
  "Coarsed exact matching" = fe_model_cem,
  "MDM matching" = fe_model_mdm
) #name all models created

cm <- c(
  "treatment_exposedTRUE:post_announcementTRUE" = "DID Interaction (Treat × Post)",
  "fake_treatment:post_val" = "Placebo: UK as Treated",
  "treatment_exposedTRUE:fake_post_2019" = "Placebo: 2019 Treatment"
) #all variables of interest

modelsummary(
  robustness_list,
  stars = c('*' = .1, '**' = .05, '***' = .01),
  coef_map = cm,
  gof_omit = 'DF|Deviance|AIC|BIC|Log.Lik|Within|Pseudo|FE',
  title = "Table X: Robustness Analysis of CSRD Impact",
  notes = "Standard errors clustered at firm (RIC) level. Controls included."
) #summary of the models
```


## Forest plot

```{r plot of all coefficients}
#clean coefficients fuunction
extract_fixest <- function(model) {
  ct <- as.data.frame(coeftable(model))
  ci <- as.data.frame(confint(model, level = 0.95))
  
  tibble(
    term = rownames(ct),
    estimate = ct[[1]],      
    conf.low = ci[[1]],      
    conf.high = ci[[2]]      
  )
}

#combine specifications list
master_plot_list <- robustness_list

#filter out the raw models, and build better categoires
df_plot_data <- map_dfr(master_plot_list, extract_fixest, .id = "specification") %>%
  filter(term %in% c(
    "treatment_exposedTRUE:post_announcementTRUE",
    "fake_treatment:post_val",
    "treatment_exposedTRUE:fake_post_2019"
  )) %>%
  mutate(
    #create the categories
    type = case_when(
      specification == "Sophisticated model" ~ "Main Model (Firm & Year FE)",
      specification %in% c("Placebo (UK)", "Placebo (T-3)") ~ "Placebo Tests",
      specification %in% c("Propensity score matching", "Coarsed exact matching", "MDM matching") ~ "Matching Extensions",
      TRUE ~ "Alternative Specifications"
    ),
    #change the clean label strings
    specification = case_when(
      specification == "Sophisticated model" ~ "Main Model (Firm & Year FE)",
      specification == "Raw GW"               ~ "Alternative DV: Raw GW Index",
      specification == "Year clustered GW"    ~ "Alternative DV: Year Standardised",
      specification == "Industry clustered GW"~ "Alternative DV: Industry Standardised",
      TRUE ~ specification
    )
  )

#construct the ggplot based on these seperate sectors, with CI and point estimates
p_stability <- ggplot(df_plot_data, aes(x = estimate, y = fct_rev(fct_inorder(specification)), color = type)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.8) +  #null line
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), size = 0.6, linewidth = 0.9) + 
  scale_color_manual(values = c(
    "Main Model (Firm & Year FE)" = "#1F77B4",
    "Alternative Specifications"  = "#2CA02C",
    "Matching Extensions"         = "#9467BD",   
    "Placebo Tests"               = "#D62728"
  )) +
  labs(
    x = "Estimated Treatment Coefficient (ATT Effect Size)",
    y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(), 
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(face = "bold", color = "black"),
    axis.title.x = element_text(margin = margin(t = 10)),
    legend.title = element_blank(),
    legend.position = "bottom",
    plot.margin = margin(15, 15, 15, 15)
  )

#save to output folder
ggsave("output/treatment_effect_stability.png", plot = p_stability, 
       width = 7.5, height = 4.8, units = "in", dpi = 300)

# print it out
print(p_stability)
```